### 2025-07-11 paper/paper_draft.tex

% !TEX root = ../paper/paper_draft.tex
\documentclass[11pt]{article} % Choose a document class, like 'article', 'report', or 'revtex4-2' for scientific papers

\usepackage{jcappub}

\usepackage[utf8]{inputenc}   % Allows UTF-8 character encoding
\usepackage{amsmath}          % For advanced math typesetting
\usepackage{amssymb}          % Additional math symbols
\usepackage{graphicx}         % For including images
\usepackage{hyperref}         % For hyperlinks in the document
\usepackage{geometry}         % For setting up page geometry
\geometry{margin=1in}         % 1-inch margins all around
\usepackage{siunitx}          % For consistent typesetting of units
\usepackage{float}            % To control float positioning (e.g., figures/tables)
\usepackage{enumitem}         % Better control over list formatting
\usepackage{tocloft}
\usepackage{tikz}
\usepackage{tikz-3dplot}
\usetikzlibrary{shapes.geometric, arrows, backgrounds}
\usepackage{empheq}
\usepackage{booktabs}
\usepackage{multirow} % needed for merging cells vertically
\usepackage{booktabs}
\usepackage[table]{xcolor}

\usepackage{listings}
\usepackage{xcolor} % For color customization (optional)

\usepackage[backend=bibtex]{biblatex}
\addbibresource{sources.bib}

\lstset{
    language=Python,
    basicstyle=\ttfamily\small,
    keywordstyle=\color{blue}\bfseries,
    stringstyle=\color{red},
    commentstyle=\color{gray}\itshape,
    showstringspaces=false,
    frame=single, % Adds a frame around the code
    breaklines=true % Wrap long lines
}

% some custom commands that should've already been in latex by default
\DeclareRobustCommand{\d}{\ifmmode\text{d}\else d\fi}
\DeclareRobustCommand{\Cov}{\ifmmode\text{Cov}\else d\fi}
\DeclareRobustCommand{\CMB}{\ifmmode\text{CMB}\else d\fi}
\DeclareRobustCommand{\gal}{\ifmmode\text{gal}\else d\fi}

\newcommand{\br}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\sbr}[1]{\ensuremath{\left[ #1 \right]}}

\setlength{\cftbeforesecskip}{5pt}

% Additional customizations (optional)
\setlength{\parindent}{0pt}   % No paragraph indentation
\setlength{\parskip}{1em}     % Add space between paragraphs
\title{Future Parameter constraints from weak lensing CMB and Galaxy Lensing Power- and Bi-spectra}

\author[1]{Jonas Frugte}
\author[1]{P. Daniel Meerburg}

\affiliation[1]{Van Swinderen Institute for Particle Physics and Gravity, University of Groningen, Nijenborgh 4, 9747 AG Groningen, The Netherlands}
% \affil[ ]{}
% \affil[ ]{\textit{E-mail:} \texttt{jonasfrugte@gmail.com}}

\emailAdd{jonasfrugte@gmail.com}
\emailAdd{p.d.meerburg@rug.nl}
\abstract{
Upcoming stage 4 surveys, such as the Simons Observatory, LSST, and Euclid are poised to measure weak gravitational lensing of the Cosmic Microwave Background (CMB) and galaxies with unprecedented precision. While the power spectrum is the standard statistic used to analyze weak lensing data, non-Gaussianity induced by the non-linear growth of structure imparts significant cosmological information into higher-order statistics. In this paper, we forecast the ability of future surveys to constrain cosmological parameters using the weak lensing power spectrum and bispectrum from both CMB and galaxy surveys, as well as their cross-correlations. We consider an eight-parameter cosmological model ($\Lambda$CDM + $\sum m_{\nu}$ + $w_{0}$) and assess parameter constraints for stage 4 survey specifications. Our analysis shows that, in the absence of systematics, both the CMB and galaxy lensing bispectra will be detectable at high signal-to-noise. We consider parameter constraints with a ``weak'' prior and with a stronger prior from the primary CMB anisotropies. In the weak prior case the lensing bispectrum substantially improves parameter constraints over using the lensing power spectrum alone, primarily by breaking parameter degeneracies. For the stronger prior we see less significant gains, particularly for the CMB lensing bispectrum. We further find that for stage 4 surveys that are able to measure lensing on small scales where non-linear structure formation becomes important, the constraining power of the lensing bispectrum can become comparable to that of the lensing power spectrum. Furthermore, we demonstrate that, both for our weak and strong priors, the synergy between CMB and galaxy lensing is substantial. Combining the two probes, particularly with both their lensing power- and bi-spectra, leads to significantly tighter constraints than achievable with either probe individually, with the greatest improvement seen for the sum of neutrino masses. Additionally, we also analyze the changes in detectability and parameter constraints when taking into account post-Born corrections to the lensing bispectra, as these were shown to be significant, especially for the CMB lensing bispectrum \cite{postborn_pratten_lewis}. We show that the conclusions of this paper remain largely the same with this correction.
}

\begin{document}

\maketitle

% \textbf{
%     \huge \noindent 
%     Parameter Forecasts From CMB $\times$ Galaxy Lensing Powerspectra $\times$ Bispectra
%     \vspace{0.5cm}
%     \\
%     \normalsize
%     Jonas Frugte, Daan Meerburg\\
%     20 February 2025
%     }\\
% \vspace{0.5cm}

%\tableofcontents

% \section{Conventions}
% Use $\mathbf k$ for wave numbers (Fourier space), $l$ for multipole moments, $m$ for magnetic quantum number. $P^{XY}(k, \eta, \eta')$ and $B^{XYZ}(k_1,k_2,k_3, \eta_1, \eta_2, \eta_3)$ correspond to bi and power spectra of fields $X$, $Y$, and $Z$ in Fourier space. $P_l^{XY}(\eta,\eta')$ and $B_{\ell_1\ell_2\ell_3}^{XYZ}(\eta_1, \eta_2, \eta_3)$ correspond to the multipole moments of the bi and power spectra of fields $X$, $Y$, and $Z$. If all times are the same we may instead write only one time argument and if the time is current time then we omit the time argument entirely. In other words,
% $$
% P_l^{XY}(\eta):=P_l^{XY}(\eta,\eta), \quad P_l^{XY}:=P_l^{XY}(\eta_{\text{current}}), \quad \text{similarly for all other spectra}.
% $$
% If both fields are the same, we may only denote one field, e.g. 
% $$
% P_l^{X}(\eta) := P_l^{XX}(\eta).
% $$

% Our Fourier convention is
% \begin{align}
%     &\text{Forward Transform:} & \quad \tilde{f}(k) &= \int_{-\infty}^{\infty} f(x) \, e^{-i k x} \, dx\\
%     &\text{Inverse Transform:} & \quad f(x) &= \frac{1}{2\pi} \int_{-\infty}^{\infty} \tilde{f}(k) \, e^{i k x} \, dk
% \end{align}
% we choose this convention because it is in line with the convention used in the CAMB source code. We work in natural units where $c=1$.

\section{Introduction}

Weak lensing has been used as a probe to constrain cosmological parameters for over a decade \cite{kilbinger2015cosmology,Bacon2000,Kaiser2000,Waerbeke2000,Bartelmann2001,Planck2018Lensing}. A range of upcoming surveys, such as the survey by the Simons Observatory (SO) \cite{Ade2019}, the Legacy Survey of Space and Time (LSST) \cite{Ivezic2019}, and Euclid \cite{Laureijs2011} aims to measure weak lensing of the cosmic microwave background (CMB) and galaxies respectively. These upcoming surveys are expected to achieve significantly higher accuracy than their predecessors, leading to tighter constraints on cosmological parameters and advancing our overall understanding of cosmology. Typically, the standard summary statistic of interest is the power spectrum (or equivalently, two-point function) of the lensing potential (as well as cross-correlations with galaxy clustering). This can be measured in terms of the lensing convergence $\kappa$, lensing shear $\gamma$, or lensing potential $\psi$. In the weak lensing regime and under the Born approximation, they are all equivalent, in the sense that they can be directly converted into one another (appendix \ref{sec:weaklensstats}). In CMB surveys, the temperature and polarization anisotropies are used to reconstruct the lensing potential \cite{Hu_2002}, while with galaxies, one instead calculates lensing effects by measuring the lensing shear, which can be deduced from the ellipticity distribution of the observed galaxies \cite{HoekstraJain2008}.

Even with Gaussian initial conditions, there are significant amounts of non-Gaussianity in the weak lensing signal, especially in galaxy lensing due to it being a probe of the more recent distribution of matter \cite{Bernardeau1997,Takada2003}. In addition to looking at the power spectrum, a natural next step is to measure weak lensing bispectra. While the bispectrum has been measured for galaxy lensing signals \cite{vanWaerbeke2002}, a definitive detection of the CMB lensing bispectrum is still pending \cite{Kalaja_2023,Namikawa_2016}. Upcoming surveys, such as the ones mentioned earlier, may be able to detect this signal. Therefore it is timely to consider the potential of higher-order spectra to constrain cosmological parameters.

Additionally, we will also look at the effects of post-Born corrections. The CMB and galaxy lensing power spectra are typically affected by less than $1\%$ \cite{postborn_pratten_lewis, galpostbornlpscorr}. The galaxy lensing bispectrum is only affected by a few percent at most \cite{galpostbornlbscorr}. The only significant effect is on the CMB lensing bispectrum, as was shown in a 2024 paper \cite{postborn_pratten_lewis}. In particular, for the folded configuration the effect can be of the same magnitude as the lensing bispectrum due to nonlinear large scale structure \eqref{eq:lensingbispectrum}. It is thus worth considering how including first order corrections changes the detectability of the CMB lensing bispectrum and the associated parameter constraints. This analysis will shed light on whether or not the the Born approximation is still appropriate to use if one were to calculate and use the CMB lensing bispectrum using future data.

In this paper, we aim to answer these questions by adopting experimental parameters similar to those of next-generation surveys. We are especially interested in seeing if approximate parameter degeneracies can be broken by combining CMB and galaxy weak lensing power- and bi-spectra. 
%Calculating the covariance matrix of unbiased optimal estimators of the cosmological parameters can be done by combining a Boltzmann code for cosmological perturbations with a Fisher matrix analysis. 
%See figure \ref{fig:flowchart} for an overview of the calculation done in this paper. 
We will consider $\Lambda$CDM parameters and extensions thereof. Of particular interest are the sum of neutrino masses $\sum m_\nu$ and dark energy equation of state parameter, $w_0$.

% While the final formula is available in the literature (e.g., the appendix of \cite{Kalaja_2021}), this paper includes a derivation as the authors were unable to locate one elsewhere.

The structure of this paper is as follows: we introduce relevant formulas for weak lensing statistics, nonlinear matter bispectrum, and Fisher matrix formalism in section \ref{sec:theory}. We specify details such as the fiducial values of the parameters and noise models used in section \ref{sec:calcdetails}. Parameter constraints and signal-to-noise ratios are presented in section \ref{sec:results} and discussed in section \ref{sec:discussion}. Useful derivations relevant to this paper can be found in the appendices. In particular, they include a derivation for the Fisher matrix formalism for the non-trivial context of bispectra with multiple tracers.

All code used to derive the results in this paper is publicly available on GitHub at \url{https://github.com/Jonas-Frugte/fisher_calc_weak_lensing}.

% \begin{figure}[h!]
%     \centering
%     \input{figures/calc_flowchart.tikz}
%     \caption{Flowchart of steps involved in calculating the minimum error/uncertainty in cosmological parameter estimates using convergence and shear power- and bi-spectra.}
%     \label{fig:flowchart}
% \end{figure}

\section{Background}
\label{sec:theory}
\subsection{Weak lensing spectra}
Radiation from cosmological objects is distorted due to gravitational lensing. Due to the low density of the cosmic web, the average deflection of a photon as it propagates through the universe is relatively weak. We thus work under the assumption that all deflection angles are small, this is referred to as the weak lensing regime. Weak lensing is quantified through the deflection field $\mathbf d(\hat {\mathbf n})$ which equals the difference between the observed angle of a point in the sky and the true (unlensed) angle. This field is the gradient of the lensing potential, $\psi(\hat {\mathbf n})$, which is expressed as a weighted integral of the mass distribution along the line-of-sight from the observer to the source.

In the case of CMB surveys, lensing alters the statistical properties of the temperature and polarization fields and can thus be calculated by comparing the observed signal to the expected unlensed signal (see e.g. \cite{cmblensingestimator, Maniyar2021_QE_CMBLensing, Hirata_2003}). In galaxy surveys, lensing alters the ellipticities of observed galaxies. If a large enough number of galaxies are observed, this effect can be separated from the intrinsic ellipticities of the galaxies which allows us to estimate the lensing potential.

To constrain cosmological parameters we can then look at the lensing potential of the CMB, $\psi_\CMB$, and of galaxy surveys, $\psi_\gal$. These are directly related to the matter power and bispectra as
\begin{align}%[box=\fbox]{align*}
    C^{\psi_X\psi_Y}_\ell
    & = \frac{9}{\ell^4}\Omega_m^2H_0^4\int_0^{\chi_*} \chi^2 \d\chi a(\eta_0-\chi)^{-2}W_X(\chi)W_Y(\chi)P^\delta(\ell/\chi,\eta_0-\chi), \label{eq:lensingpowerspectrum}\\
    B^{\psi_X\psi_Y\psi_Z}_{\ell\ell_1\ell\ell_2\ell\ell_3} &= \sqrt{\frac{(2\ell_1 + 1)(2\ell_2 + 1)(2\ell_3 + 1)}{4\pi}} \begin{pmatrix} \ell_1 & \ell_2 & \ell_3 \\ 0 & 0 & 0 \end{pmatrix} \frac{27}{\ell_1^2\ell_2^2\ell_3^2}\Omega_m^3H_0^6 \notag\\
    & \quad \times \int \chi^2\d \chi a(\eta_0-\chi)^{-3}W_X(\chi)W_Y(\chi)W_Z(\chi)  B^\delta(\{\ell_i/\chi\}, \eta_0-\chi), \label{eq:lensingbispectrum}
\end{align}
with $X, Y, Z \in \{\CMB, \gal\}$. A derivation can be found in appendices \ref{sec:weaklensing} and \ref{sec:weaklensstats}, or in the literature; see, e.g. \cite{bartelmann2001weak}. $\Omega_m$ is the present day matter density parameter. $H_0$ is the present-day Hubble constant, $a(\eta)$ is the scale factor, $\eta_0$ is the conformal time today and $\chi$ is the comoving radial distance. $\chi_*$ presents the distance to surface of last scattering. $P^\delta(k, \eta)$ is the matter power spectrum. $B^\delta(k_1, k_2, k_3, \eta)$ is the matter bispectrum. $W_X(\chi)$ is the window function. Finally,
$$
\begin{pmatrix}
    \ell_1&\ell_2&\ell_3 \\ m_1 & m_2 & m_3
\end{pmatrix},
$$ 
is the Wigner 3-j symbol. The difference between CMB and galaxy lensing is the window function, defined as
\begin{equation}
    W_X(\chi) = \int_\chi^{\infty}\d\chi' p_X(\chi')\frac{\chi'-\chi}{\chi'\chi},
    \label{eq:windowfuncdef}
\end{equation}
with $p_X$ the radial distribution of the radiation source. For the CMB we take $p(\chi) = \delta(\chi - \chi_*)$. The window function presents the distribution of redshifts at which a CMB photon or galaxy photon is deflected. As a result, the CMB lensing window function is broader and peaks at higher redshift compared to the galaxy window function.

Our convention for defining the linear and non-linear matter power spectrum is
\begin{equation*}
    \langle \delta(\mathbf k, \eta) \delta(\mathbf{k}',\eta) \rangle = (2\pi)^3\delta_D(\mathbf k + \mathbf k ')P^\delta(\mathbf k, \eta),
\end{equation*}
and for the bispectrum it is 
\begin{equation*}
    \langle \delta(\mathbf k_1, \eta) \delta(\mathbf k_2, \eta)\delta(\mathbf k_3, \eta) \rangle = (2\pi)^3\delta_D(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\delta(k_1,k_2,k_3, \eta),
\end{equation*}
with $\delta$ the fractional matter density perturbation field and $\delta_D$ the Dirac delta function.

\subsection{Nonlinear matter bispectrum}

Determining the nonlinear matter power spectrum can be done using numerical codes such as CAMB \cite{Lewis2000}. The nonlinear matter bispectrum was calculated from the power spectrum using a fitting formula based on perturbation theory in \cite{bispfit}. It is given by
\begin{equation}
    B^\delta(k_1, k_2, k_3, \chi) = 2 F_2(k_1, k_2, z) P^\delta(k_1, z) P^\delta(k_2, z) + \text{2 perm},
\end{equation}
where \( P^\delta \) is the nonlinear matter power spectrum\footnote{Compare to the tree level bispectrum where we instead use the linear power spectrum.}, and the kernel \( F_2 \) is modified from the tree level result with factors \( a(k, z) \), \( b(k, z) \), and \( c(k, z) \):
\begin{equation}
    F_2(k_1, k_2, z) = \frac{5}{7} a(k_1, z) a(k_2, z) + \frac{k_1^2 + k_2^2}{2 k_1 k_2} b(k_1, z) b(k_2, z) \cos \theta + \frac{2}{7} c(k_1, z) c(k_2, z) \cos^2 \theta.
\end{equation}
They are defined as:
\begin{equation}
    a(k, z) = \frac{1 + \sigma_8^{a_6}(z) \sqrt{0.7} Q(n_{\text{eff}}) (q^{a_1})^{n_{\text{eff}} + a_2}}{1 + (q^{a_1})^{n_{\text{eff}} + a_2}},
\end{equation}
\begin{equation}
    b(k, z) = \frac{1 + 0.2 a_3 (n_{\text{eff}} + 3) (q^{a_7})^{n_{\text{eff}} + 3 + a_8}}{1 + (q^{a_5})^{n_{\text{eff}} + 3.5 + a_8}},
\end{equation}
\begin{equation}
    c(k, z) = \frac{1 + \left[ \frac{4.5 a_4}{1.5 + (n_{\text{eff}} + 3)^4} \right] (q^{a_5})^{n_{\text{eff}} + 3 + a_9}}{1 + (q^{a_5})^{n_{\text{eff}} + 3.5 + a_9}}.
\end{equation}
Here, \( Q(n_{\text{eff}}) \) is given by:
\begin{equation}
    Q(x) = \frac{4 - 2x}{1 + 2^{x+1}}.
\end{equation}
The effective spectral index of the linear power spectrum is defined as:
\begin{equation}
    n_{\text{eff}} \equiv \frac{d \ln P^\delta_{\text{lin}}(k)}{d \ln k}.
\end{equation}
Additionally, \( q \) is given by:
\begin{equation}
    q = \frac{k}{k_{\text{NL}}},
\end{equation}
where \( k_{\text{NL}} \) is the scale at which nonlinearities become significant, satisfying:
\begin{equation}
    4 \pi k_{\text{NL}}^3 P^\delta_{\text{lin}}(k_{\text{NL}}, 0) = 1.
\end{equation}
The coefficients \( a_i \) are:
\begin{align*}
    a_1 &= 0.484, \quad a_2 = 3.740, \quad a_3 = -0.849, \quad a_4 = 0.392, \\
    a_5 &= 1.013, \quad a_6 = -0.575, \quad a_7 = 0.128, \quad a_8 = -0.722, \quad a_9 = -0.926.
\end{align*}

\subsection{Post-Born corrections}
One of the approximations made in the derivation of the weak lensing power- and bispectra (equations \eqref{eq:lensingpowerspectrum} and \eqref{eq:lensingbispectrum}) is that the integral is performed along the line of sight to the light source. To get an analytically correct result, one would instead need to solve a differential equation with a feedback loop where the effects of gravitational lensing affect the trajectory and shape of given bundle of light rays which in turn changes the way future gravitational lenses affect the bundle. One can also choose to make order by order corrections to the Born approximation. There is motivation to understand the effects of these corrections (see the introduction) so we present them here.

Under the Born approximation, the elements of the distortion tensor, $\psi_{ab}$, (from which shear, convergence, and rotation are constructed) equal second derivatives of the lensing potential. We thus calculate the potential instead of the distortion tensor itself. When post-born effects are taken into account this is no longer true and we are forced to work with $\psi_{ab}$ directly. Following \cite{postborn_pratten_lewis}, the leading order post born correction to the distortion tensor for a point source at distance $\chi_s$ is given by
\begin{equation}
    \left( \psi_{ab}^{\text{p.s.}}  \right)^{(2)}(\chi_s) := 2 \int^{\infty}_0 d \chi^{} \, \chi^2 \, \frac{\chi_s - \chi}{\chi_s\chi}\Theta(\chi_s - \chi) \, \left[ - \Psi_{,ac} (\chi) \left(\psi^{\text{p.s.}}_{cb}\right)^{(1)} (\chi) + \Psi_{,acd} (\chi) \, \delta x_d^{(1)} (\chi^{}) \right].
    \label{eq:pbcorrectionpointsorce}
\end{equation}
Here $\left(\psi_{cb}^{(p.s.)}\right)^{(1)}(\chi)$ is the deformation for a point source at $\chi$ under the Born approximation as defined earlier. $\delta x_d^{(1)}(\chi)$ is the first order correction to the location of the lensed light ray at a distance $\chi$, given by
\begin{equation}
\delta x_a^{(1)} = -2 \int^{\chi}_0 d \chi^{\prime} \, \frac{\chi-\chi'}{\chi\chi'} \, \chi \chi' \, \Psi_{,a} (\chi').
\end{equation}
To adapt the above to a source with radial distribution $p(\chi)$ (which in our case will be a population of galaxies) we integrate over $\chi_s$ as follows:
\begin{equation}
    (\psi_{ab}^{\text{gal}})^{(i)} = \int_0^\infty d \chi_s p(\chi_s)\left(\psi_{cb}^{(p.s.)}\right)^{(i)}(\chi_s).
\end{equation}
This step essentially just changes $(\chi_s - \chi)/(\chi_s\chi)\Theta(\chi_s-\chi)$ in equation \eqref{eq:pbcorrectionpointsorce} to the window function $W(\chi)$ defined in equation \eqref{eq:windowfuncdef}.

The leading order correction to the convergence bispectrum of sources 1, 2, 3 is given as
\begin{equation}
    B^{\kappa_{(1)}\kappa_{(2)}\kappa_{(3)}}_{L_1 L_2 L_3} = 2\frac{\mathbf L_1\cdot\mathbf L_2}{L_1^2L_2^2}\left[ \mathbf L_1\cdot \mathbf L_3 \mathcal M_s^{(1)(2)(3)}(L_1,L_2)
 +\mathbf L_2\cdot \mathbf L_3 \mathcal M_s^{(2)(1)(3)}(L_2,L_1)\right] + \text{cyc. perm.},
\end{equation}
where
\begin{equation}
    \mathcal M_s^{(1)(2)(3)}(L_1,L_2) := L_1^4\int_0^{\chi_s}d\chi \frac{W^{(1)}(\chi)W^{(3)}(\chi)}{\chi^2} P_{\Psi\Psi}\left(\frac{L_1}{\chi}, z(\chi)\right) C_{L_2}^{\kappa_{(2)}\kappa_{(3)}}(\chi,\chi_s),
\end{equation}
and
\begin{equation}
    C_{L_2}^{\kappa_{(2)}\kappa_{(3)}}(\chi,\chi_s) := \int_0^{\infty}d \chi' W^{(2)}(\chi')\frac{\chi_s-\chi}{\chi_s\chi}\Theta(\chi_s-\chi)P_{\Psi\Psi}(\frac{L_2}{\chi'},\chi').
\end{equation}

\subsection{Fisher matrix analysis}
The signal-to-noise (SNR) and parameter constraints are obtained through a Fisher matrix analysis as is standard in cosmology\cite{dodelson2020modern}. A full derivation of the equations used below can be found in appendix \ref{sec:fisher}. The auto- and cross- power spectrum Fisher matrix is given as
\begin{equation*}
    F_{\alpha\beta} = \sum_{\ell} \sum_{XY}\sum_{X'Y'}(2\ell+1)\partial_\alpha C_\ell^{XY} (C^{-1})_\ell^{XX'}(C^{-1})_\ell^{YY'} \partial_\beta C^{X'Y'}_\ell,
\end{equation*}
where
\begin{equation*}
    C_{\ell} := \begin{pmatrix}
        C^{\psi_{\text{CMB}}\psi_{\text{CMB}}}_\ell & C^{\psi_{\text{CMB}}\psi_{\text{gal}}}_\ell \\
        C^{\psi_{\text{CMB}}\psi_{\text{gal}}}_\ell & C^{\psi_{\text{gal}}\psi_{\text{gal}}}_\ell
    \end{pmatrix}.
\end{equation*}
For the bispectra we instead have 
\begin{equation*}
    F_{\alpha\beta} = \sum_{\ell_1 \leq \ell_2 \leq \ell_3} \frac{\mathcal P _{\ell_1\ell_2\ell_3}}{6}
    \sum_{XYZ}\sum_{X'Y'Z'} 
    \partial_\alpha B^{X Y Z}_{\ell_1 \ell_2 \ell_3} 
    (C^{-1})^{X X'}_{\ell_1}
    (C^{-1})^{Y Y'}_{\ell_2}
    (C^{-1})^{Z Z'}_{\ell_3}
    \partial_\beta B^{X' Y' Z'}_{\ell_1 \ell_2 \ell_3},
\end{equation*}
with $\mathcal P_{\ell_1\ell_2\ell_3}$ defined as the number of distinct permutations that can be made with $\ell_1\ell_2\ell_3$. When only considering auto spectra the formulas simplify to:
\begin{equation*}
    F_{\alpha\beta} = \sum_{\ell} \sum_{XY}\sum_{X'Y'}\frac{2\ell+1}{2}\frac{\partial_{\alpha}C_\ell^{XX} \partial_\beta C^{XX}_\ell}{(C_\ell^{XX})^2},
\end{equation*}
and (for bispectra, assuming Gaussian, diagonal, covariance)
\begin{equation*}
    F_{\alpha\beta} = \sum_{\ell_1 \leq \ell_2 \leq \ell_3} \frac{\mathcal P _{\ell_1\ell_2\ell_3}}{6}
    \frac{\partial_\alpha B^{X X X}_{\ell_1 \ell_2 \ell_3} 
    \partial_\beta B^{XXX}_{\ell_1 \ell_2 \ell_3}}{C^{XX}_{\ell_1}C^{XX}_{\ell_2}C^{XX}_{\ell_3}}.
\end{equation*}
If only some fraction of the sky, $f_{\text{sky}}$, is measured, all Fisher matrices are multiplied by $f_{\text{sky}}$ \cite{Takada2003}.

\section{Experimental parameters and priors}\label{sec:calcdetails}
% TODO: find proper references for stage 3 and stage 4 names as well as experimental parameters and real life surveys
\subsection{Fiducial cosmology}
Our fiducial cosmology is based on the Planck results \cite{planckresults} (see table \ref{tab:fiducialpars}). We choose to constrain the standard parameters of the $\Lambda$CDM model as well as neutrino mass, $m_\nu$, and dark energy equation of state parameter, $w_0$. Our reason for including these extra degrees of freedom is that we would like to see to what degree future surveys will be able to tell us more about the mass of neutrinos and the nature of dark energy. For example, a paper from the DESI collaboration \cite{Roy2024DESI} recently showed that measurements of baryon acoustic oscillations give a 2.5 - 4 $\sigma$ tension with a non-evolving dark energy equation of state model, so it is worth considering if future weak lensing experiments could tell us more about this tension. Regarding neutrinos, depending on how stage 4 surveys will be able to constrain neutrino mass, the results could tell us more about their mass hierarchy and serve as a test for constraints obtained from other probes, such as baryon acoustic oscillations.

\subsection{Choice of priors}
Throughout this paper we consider two priors. The first we refer to as our ``weak prior'' and is based on the one used in \cite{Planck2018Lensing} (see table \ref{tab:fiducialpars}). This prior only restricts $\Omega_b h^2$ and $n_s$ significantly, as those are not well constrained by lensing spectra alone. Our differences with \cite{Planck2018Lensing} are that (i) we do not fix $\tau$, but instead take it to have a SNR of 1 and added a loose $A_s$ prior as well, (ii) instead of a flat prior for the Hubble constant ($40 < H_0 < 100$) we use a gaussian distribution with the same standard deviation, (iii) we added a weak constraint on $\Omega_c h^2$, and (iv) we also vary neutrino mass and $w_0$ but do not assume any constraints on them in our priors. The purpose of the weak prior is as follows. The weak lensing power spectra, in particular, exhibit complete insensitivity to certain parameters. Consequently, relying solely on them to constrain the entire cosmological model results in very poor constraints. Future lensing surveys will of course use information from surveys measuring other statistics such as those derived from galaxy clustering and the primary CMB anisotropies. By adding a prior based on these (primary) statistics our results become more representative. Additionally, by considering a conservative prior, it remains clear how the different lensing spectra complement each other. The other prior we consider is based on the CMB temperature and $E$-mode polarization cross- and auto-powerspectra from $\ell=30$ to $\ell = 2000$ assuming the same noise properties as used for the CMB weak lensing reconstruction noise. Adopting these priors  serve to test how weak lensing statistics are able to further tighten constraints beyond measuring primary statistics. It is worth noting that the constraint for $\tau$ and $w_0$ are tighter than in other literature (see \cite{cmbs4sciencebook} and \cite{Namikawa_2016} for comparable $\tau$ and $w_0$ constraints, respectively). Based on a number of internal tests\footnote{
This included looking at the accuracy of numerical derivatives w.r.t. cosmological parameters, testing the numerical stability of Fisher matrices wr.r.t. matrix inversion, and varying our values for detector noise and beam width.
} we suspect that this is mainly due us not taking into account foregrounds in our noise models.

\begin{table}[h!]
    \centering
    \scriptsize
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        \textbf{parameter} & \textbf{notation} & \textbf{fiducial value} & \textbf{$\sigma$ of weak prior} & \textbf{$\sigma$ of $T + E$ prior}\\
        \hline
        Hubble constant & $H_0$ & $67.4 \, \text{km/s/Mpc}$ & 17.3 & 1.21\\
        \hline
        Physical baryon density parameter & $\Omega_b h^2$ & $0.0223$ & 0.0005 & 0.000057\\
        \hline
        Physical cold dark matter density parameter & $\Omega_c h^2$ & $0.119$ & $0.288$ & 0.00083\\
        \hline
        Scalar spectral index & $n_s$ & 0.965 & 0.02 & 0.0025 \\
        \hline
        Reionization depth & $\tau$ & 0.063 & 0.063 & 0.0013 \\
        \hline
        Amplitude of primordial scalar fluctuations & $A_s$ & $2.13 \times 10^{-9}$ & $1\times 10^{-9}$ & $5.5\times 10^{-11}$\\
        \hline
        Sum of neutrino masses & $\sum m_\nu$ & $0.06$ & $\infty$ & 0.22\\
        \hline
        Dark energy equation of state parameter & $w_0$ & -1 & $\infty$ & 0.058 \\ 
        \hline
    \end{tabular}
    \caption{Cosmological parameters varied ($\Lambda \text{CDM} + w_0 + m_\nu$). We use a prior similar to the one used by the Planck weak lensing results \cite{Planck2018Lensing} which essentially only restricts $\Omega_bh^2$ and $n_s$ significantly.}
    \label{tab:fiducialpars}
\end{table}

\subsection{Noise Modeling}
This paper considers noise levels for ``stage 3'' and ``stage 4'' (weak lensing) surveys. Noise properties considered for this analysis can be found in table \ref{tab:noiselevels}, including experiments they are representative of. A comparison of the noise power spectra to the lensing power spectra are shown     in figure \ref{fig:lpsplusnoise}. Effectively, for galaxy lensing, the power spectrum is signal-dominated up until $\ell \sim 200$ (stage 3) and $\ell \sim 700$ (stage 4). For CMB lensing, the power spectrum is signal-dominated for $\ell \sim 300$ (stage 3) and $\ell \sim 1000$ (stage 4).

\begin{table}[h!]
    \centering
    \begin{tabular}{|c|c|c|c|}
    \hline
    \textbf{source} & \textbf{survey stage} & \textbf{noise vals} & \textbf{comparable experiments} \\
    \hline 
    \multirow{2}{*}{CMB} 
        & stage 3 & $\sigma = 1'$, $\Delta_P = 6' \; \mu \text{K}$ & Advanced ACTPol, Simons Observatory \\
    \cline{2-4}
        & stage 4 & $\sigma = 3'$, $\Delta_P = 1' \; \mu \text{K}$ & CMB-S4 \\
    \hline
    \multirow{2}{*}{galaxies} 
        & stage 3 & $\sigma_{\text{rms}} = 0.3$, $n_g = 5 \text{arcmin}^{-2}$ & DES, KiDS \\
    \cline{2-4}
        & stage 4 & $\sigma_{\text{rms}} = 0.3$, $n_g = 30 \text{arcmin}^{-2}$ & LSST, Euclid \\
    \hline
    \end{tabular}
    \caption{Noise levels considered for weak lensing of galaxies and the CMB. $\sigma$ (beam width) and $\Delta_P$ (polarization white noise) describe CMB survey specifications, while $\sigma_{\text{rms}}$ (intrinsic galaxy ellipticity) and $n_g$ (observed galaxy density) refer to galaxy shear surveys.}
    \label{tab:noiselevels}
\end{table}

The CMB lensing noise is estimated using a quadratic estimator \cite{cmblensingestimator}.
% We consider experiments that only measure polarization noise. This is similarly done in \cite{Namikawa_2016}. 
The parameters characterizing the noise levels are beam width, $\sigma$, polarization white noise, $\Delta_P$, and temperature white noise $\Delta_T$. We take $\Delta_T = \Delta_P / \sqrt 2$ throughout. Galaxy lensing is determined by measuring lensing shear (see appendices \ref{sec:weaklensing} and \ref{sec:shear}). The noise in this measurement is dominated by scale-independent shot noise and has associated noise power spectrum $N_l^{\text{shear}} = \sigma_{\text{rms}}^2 / n_g$, \cite{bartelmann2001weak}. $n_g$ is the amount of galaxies observed per unit solid angle. The noise power spectrum for the lensing potential then equals
\begin{equation*}
    N_\ell^{\text{lens. potential}} = \frac{4}{(\ell-1)\ell(\ell+1)(\ell+2)}N_\ell^{\text{shear}}.
\end{equation*}
In all cases we assume that the proportion of the sky surveyed, $f_{\text{sky}}$, equals 0.5.
For parameter constraints, we will consider only stage 4 noise levels.

\begin{figure}[t]
    \centering
    \includegraphics[width=\textwidth]{../code/plots/spectraplusnoise.pdf}
    \caption{CMB (right) and galaxy (left) lensing potential power spectra compared to associated experimental noise. Current (stage 3) noise values are displayed as well as near future\footnote{Some of the stage 4 experiments are already in operation, but expected noise levels are only achieved after several years of integration.} (stage 4) noise values. The CMB lensing experiment uses only polarization. CMB noise values are chosen in accordance with \cite{Namikawa_2016}. For comparison we also how the reconstruction noise for the Simons Observatory. Shear noise values are chosen to be similar to e.g. Euclid measurements for stage 4 and e.g. KiDS for stage 3.}
    \label{fig:lpsplusnoise}
\end{figure}

% For noise in CMB lensing we consider lensing estimated by an experiment measuring polarization with white noise $\Delta_P$ and beam spread $\sigma$. We then use the quadratic estimator from \cite{Hu_2002} to estimate the resulting noise in the lensing power spectrum.

% For shear measurements, the noise is simpler to model as the primary source is shot noise due to the discrete number of galaxies in each patch of the survey. The mean ellipiticity of the galaxies in a patch is calculated and a deviation from the mean is interpreted as lensing shear. The greater the number of galaxies the more accurately can the mean ellipticity be determined. Choosing the size of the patch is a balancing between signal to noise and resolution. Smaller patches give more uncertain shear measurements while providing better overal resolution for the shear map and thus allowing measurements of shear bi- and powerspectra to higher multipole numbers. The opposite is true for larger patch sizes. The power spectrum of the shear noise is given by
% \begin{equation*}
%     N^{\gamma\gamma}_{l} = \frac{\sigma^2_{rms}}{\bar n_g}
% \end{equation*}
% where $\bar n_g$ is the mean number of galaxies observed per sterradian accross the whole survey \cite{weaklensingnotes2020}.
% and $\sigma^2_\gamma$ is the intrinsic ellipticity dispersion. 
% Experiments such as Euclid and the LSST both have $\sigma_{rms} \approx 0.3$ \cite{euclid1overview} \cite{lsstsciencebook} and 
% $ n_g\approx 30 \text{ arcmin}^{-2}$.

% The noise values are calculated for stage 3 (current) surveys and stage 4 (near future) surveys. We will use the stage 4 noise values for parameter constraints. A comparison of stage 3 noise, stage 4 noise, and lensing signal can be seen in figure \ref{fig:lpsplusnoise}.
\subsection{Other details}
The redshift distribution of the observed galaxies is commonly parameterized as \cite{bartelmann2001weak}
\begin{equation*}
    n(z) \propto z^a \exp\sbr{-\br{\frac{z}{z_0}}^b}.
\end{equation*}
We choose the parameter combination $a = 2$, $b = 3/2$ and $z_0 = 0.64$ which is similar to the expected distributions of Euclid, (which will probe primarily in the 0.2 - 2.6 redshift range \cite{euclidprep10})
and the LSST mission (which has $a \approx 2$, $b \approx 1$, and $z_0 \approx 0.3$ from predictions for the obtained data \cite{lsstsciencebookchapter3}). For simplicity, we do not implement a tomographic binning of the source galaxies, and instead model the population with a single effective redshift distribution.

Derivatives are calculated with a central difference formula (see appendix \ref{sec:derivatives} for convergence).  To check the accuracy of our results we calculated constraints for similar experimental parameters as in references \cite{Planck2018Lensing, Takada2003, cmbs4sciencebook,neutrinoconstraints} and found good agreement in all cases.

\section{Results}\label{sec:results}
\subsection{Detectability}
As expected, lensing power spectra should be detectable at high signal-to-noise. 
The signal-to-noise of the CMB and galaxy lensing bispectra versus the maximum multipole moment measured is shown in figure \ref{fig:snrplots}. In the absence of systematics, shear bispectra can be detected at high signal-to-noise with both stage 3 and stage 4 experiments. CMB lensing bispectra could be  detectable by a stage 3 experiment (see e.g. ref. \cite{Kalaja_2023} for the effect of noise biases) and with high signal-to-noise with a stage 4 experiment. Our results for CMB weak lensing $S/N$ match those of reference \cite{Namikawa_2016}.

Regarding post-Born corrections, we found that the effect on the galaxy lensing bispectrum $S/N$ is a few percent at most, in agreement with \cite{galpostbornlbscorr}. We thus did not include additional plots for these corrections. For the CMB the $S/N$ is reduced significantly (in agreement with \cite{postborn_pratten_lewis}) but not so severely that they are no longer detectable. Our conclusions thus remain the same: the CMB bispectrum is probably detectable with stage 3 surveys and definitely detectable with stage 4 surveys.

\begin{figure}[h!]
    \includegraphics[width=\textwidth]{figures/snrplots.pdf}
    \caption{Signal-to-noise ratios for galaxy lensing (left) and CMB lensing (right) bispectra for stage 3 (dashed) and stage 4 (not dashed) surveys as a function of maximum multipole measured. The minimum multipole is always $l_{\text{min}}=2$. The multipole range used for CMB lensing reconstruction is $2\leq \ell \leq 10^4$. Using a more conservative range such as $30\leq \ell \leq 5000$ used to calculate Simons Observatory noise curves don't significantly affect the results.}
    \label{fig:snrplots}
\end{figure}

\subsection{Parameter Constraints}
Next, we consider cosmological parameter constraints. We include both the power spectrum and the bispectrum of the weak lensing. We look at two extension of $\Lambda CDM$, $m_\nu$ and $w_0$ as well as the derived parameters $\sigma_8$, and $\Omega_m$. Constraints on the full set of $\Lambda$CDM parameters can be found in appendix \ref{sec:lambdacdmconstraints}. Results shown include marginalization over all parameters. Figures \ref{fig:paramconstraintstightcmb} and \ref{fig:paramconstraintstightgal} show confidence ellipses assuming a weak prior from stage 4 CMB and galaxy lensing surveys, respectively. Figures \ref{fig:paramconstraintstightcmbcmbprior} and \ref{fig:paramconstraintstightgalcmbprior} similarly show confidence ellipses in the case of our primary CMB $T+E$ prior. Table \ref{tab:paramconstraintstight} summarizes all forecasted constraints.

Constraints are significantly improved by including bispectrum information in nearly all cases when using weak priors. When including the CMB $T+E$ priors, both the CMB lensing power- and bispectra significantly improve prior constraints, however combining them does not lead to significant improvements compared to only using lensing power spectra. For galaxy lensing, the power- and bispectra yield nearly identical improvements compared to the CMB $T+E$ prior. In this case, combining both statistics leads to significantly tighter constraints (in contrast to CMB lensing). When combining CMB and galaxy lensing we see that the bispectrum is slightly less competitive compared to the lensing power spectrum, but can still significantly improve constraints compared to only using the lensing power spectrum.
% Again in agreement with reference \cite{Namikawa_2016}, in which case a 33 percent and 35 percent improvement was found for $m_\nu$ and $w_0$, respectively, by measuring the CMB lensing bispectrum in addition to the power spectrum.

Furthermore, it is clear that, in general, combining CMB and galaxy lensing surveys also leads to significant improvements in constraining power, even in the case of a strong CMB $T+E$ prior. The improvement is most significant for the neutrino mass constraint.

Besides these obvious findings, we make several other conclusions:
\begin{itemize}
    \item Lensing power spectra alone are not sufficient to competitively constrain $\Lambda\mathrm{CDM} + w_0 + m_\nu$ without priors, even in the case of stage 4 surveys.

    \item For lower noise values (in particular for stage 4 noise), the bispectra become increasingly important compared to the power spectra. The lensing bispectra are directly related to the amount of non-Gaussianity in the matter distribution. As noise levels become lower, we are able to measure the matter distribution on small enough scales where nonlinear matter evolution becomes important. Additionally, for galaxy lensing, the window function peaks at later redshifts compared to CMB lensing. The bispectrum is largest at late times, and thus the galaxy lensing bispectrum is significantly easier to detect and is better at constraining parameters than the CMB lensing bispectrum.

    \item The information from bispectra is sensitive to cosmology in a different way than that of the power spectra. The main parameter combinations where approximate degeneracies are broken are $(n_s,\Omega_b h^2),\;(\Omega_b h^2,H_0),\;(\Omega_c h^2,H_0),\;(\Omega_c h^2,\Omega_b h^2)\;(n_s,\Omega_c h^2),\;(m_\nu,H_0),\;(m_\nu,n_s),$ $(A_s,\Omega_b h^2)$ (see also appendix \ref{sec:lambdacdmconstraints}). This means that even if stage 4 surveys are not able to reach the noise levels assumed in this paper, the bispectra will likely still lead to significantly tighter constraints.

    \item Stage 4 galaxy weak lensing surveys generally contain more information than CMB surveys. 
    Comparing the confidence ellipses, it is clear that the two types of surveys depend on the underlying parameters in different ways. This leads to substantially better constraints for $\sigma_8$, $\Omega_m$, $m_\nu$, and $w_0$ when combining CMB and galaxy lensing.

    \item It is clear that post-Born corrections do not change the constraints drastically. As expected, they matter less for galaxy lensing and when using the CMB $T+E$ prior. Interestingly, although the constraints due to bispectra alone are always worse, for some parameters, they are instead improved when combining lensing bispectrum information with that of the lensing power spectrum. This implies that, at least in some cases, the post-Born contribution to the lensing bispectrum depends on the parameters in a different way than the lensing power spectrum and thus improves the amount of information we can extract from survey data despite making the lensing bispectrum less detectable. Besides these observations, we conclude that taking into account post-Born corrections does not lead to different conclusions about the overall sensitivity of the lensing spectra.
\end{itemize}

\begin{table}
    \centering
    \scriptsize
%     \begin{tabular}{|l|r|rrr|rrr|rrr|}
%         \hline
%         \multicolumn{11}{|c|}{weak priors} \\
%         \hline
%         && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
%         \hline
%         Par            &   prior &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
% \hline
%  $m_\nu$    & 1.000e+01 &  2.815e-01 &                  1.104e+00 &                           2.247e-01 &  9.954e-01 &                  2.646e-01 &                           1.551e-01 &  1.042e-01 &                  1.317e-01 &                           6.537e-02 \\
%  $w_0$      & 1.000e+01 &  5.415e-01 &                  1.405e+00 &                           3.683e-01 &  2.638e-01 &                  7.817e-02 &                           2.475e-02 &  5.936e-02 &                  6.629e-02 &                           1.899e-02 \\
%  $\sigma_8$ & 2.457e+00 &  1.181e-01 &                  1.872e-01 &                           5.486e-02 &  8.069e-02 &                  3.668e-02 &                           1.299e-02 &  1.375e-02 &                  1.587e-02 &                           8.048e-03 \\
%  $\Omega_m$ & 6.702e-01 &  1.948e-01 &                  1.790e-01 &                           7.270e-02 &  1.434e-01 &                  4.831e-02 &                           1.667e-02 &  1.619e-02 &                  1.607e-02 &                           7.233e-03 \\
%        \hline
%         \multicolumn{11}{|c|}{CMB $T + E$ priors} \\
%         \hline
%         && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
%         \hline
%          Par            &   prior &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
%         \hline
%  $m_\nu$    & 2.157e-01 &  7.445e-02 &                  1.178e-01 &                           7.283e-02 &  1.406e-01 &                  1.042e-01 &                           8.592e-02 &  4.340e-02 &                  8.965e-02 &                           3.735e-02 \\
%  $w_0$      & 5.781e-02 &  4.606e-02 &                  5.285e-02 &                           4.578e-02 &  4.756e-02 &                  3.934e-02 &                           2.080e-02 &  3.231e-02 &                  3.667e-02 &                           1.741e-02 \\
%  $\sigma_8$ & 1.547e-02 &  1.360e-02 &                  1.492e-02 &                           1.349e-02 &  1.252e-02 &                  1.136e-02 &                           7.881e-03 &  9.682e-03 &                  1.023e-02 &                           6.126e-03 \\
%  $\Omega_m$ & 1.495e-02 &  1.381e-02 &                  1.457e-02 &                           1.376e-02 &  1.336e-02 &                  1.124e-02 &                           8.715e-03 &  1.028e-02 &                  9.616e-03 &                           5.484e-03 \\
%        \hline
%             \multicolumn{11}{|c|}{weak priors, with post-Born corrections} \\
%         \hline
%         && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
%         \hline
%         Par            &   prior &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
% \hline
% $m_\nu$    & 1.000e+01 &  2.815e-01 &                  1.158e+00 &                           1.805e-01 &  9.954e-01 &                  3.349e-01 &                           2.073e-01 &  1.042e-01 &                  1.775e-01 &                           6.256e-02 \\
%  $w_0$      & 1.000e+01 &  5.415e-01 &                  8.193e-01 &                           3.051e-01 &  2.638e-01 &                  8.328e-02 &                           2.825e-02 &  5.936e-02 &                  6.271e-02 &                           1.988e-02 \\
%  $\sigma_8$ & 2.457e+00 &  1.181e-01 &                  1.846e-01 &                           5.660e-02 &  8.069e-02 &                  4.047e-02 &                           1.317e-02 &  1.375e-02 &                  1.782e-02 &                           8.369e-03 \\
%  $\Omega_m$ & 6.702e-01 &  1.948e-01 &                  1.810e-01 &                           7.086e-02 &  1.434e-01 &                  5.474e-02 &                           1.811e-02 &  1.619e-02 &                  1.799e-02 &                           7.325e-03 \\
%        \hline
%         \multicolumn{11}{|c|}{CMB $T + E$ priors, with post-Born corrections} \\
%         \hline
%         && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
%         \hline
%          Par            &   prior &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
%         \hline
%  $m_\nu$    & 2.157e-01 &  7.445e-02 &                  1.298e-01 &                           7.313e-02 &  1.406e-01 &                  1.166e-01 &                           9.869e-02 &  4.340e-02 &                  1.034e-01 &                           3.766e-02 \\
%  $w_0$      & 5.781e-02 &  4.606e-02 &                  5.308e-02 &                           4.596e-02 &  4.756e-02 &                  4.163e-02 &                           2.220e-02 &  3.231e-02 &                  3.911e-02 &                           1.798e-02 \\
%  $\sigma_8$ & 1.547e-02 &  1.360e-02 &                  1.498e-02 &                           1.354e-02 &  1.252e-02 &                  1.168e-02 &                           7.717e-03 &  9.682e-03 &                  1.073e-02 &                           6.285e-03 \\
%  $\Omega_m$ & 1.495e-02 &  1.381e-02 &                  1.467e-02 &                           1.379e-02 &  1.336e-02 &                  1.141e-02 &                           8.530e-03 &  1.028e-02 &                  1.017e-02 &                           5.687e-03 \\
%        \hline
%         \end{tabular}
 \begin{tabular}{|l|r|rrr|rrr|rrr|}
        \hline
        \multicolumn{11}{|c|}{weak priors (with and without post-Born corrections)} \\
        \hline
        && \multicolumn{3}{c|}{CMB lensing} & \multicolumn{3}{c|}{Gal. lensing} & \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
        \hline
        Par & prior & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
        \hline
$m_\nu$    & 10 & 0.28 & 1.1 & 0.23 & 1.0 & 0.26 & 0.16 & 0.10 & 0.13 & 0.065 \\
\rowcolor{blue!10}
           &    &  & 1.2 & 0.18 & & 0.33 & 0.21 & & 0.18 & 0.063 \\
$w_0$      & 10 & 0.54 & 1.4 & 0.37 & 0.26 & 0.078 & 0.025 & 0.059 & 0.066 & 0.019 \\
\rowcolor{blue!10}
           &    &  & 0.82 & 0.31 &  & 0.083 & 0.028 & & 0.063 & 0.020 \\
$\sigma_8$ & 2.5 & 0.12 & 0.19 & 0.055 & 0.081 & 0.037 & 0.013 & 0.014 & 0.016 & 0.0080 \\
\rowcolor{blue!10}
           &     &  & 0.18 & 0.057 & & 0.040 & 0.013 & & 0.018 & 0.0084 \\
$\Omega_m$ & 0.67 & 0.19 & 0.18 & 0.073 & 0.14 & 0.048 & 0.017 & 0.016 & 0.016 & 0.0072 \\
\rowcolor{blue!10}
           &     &  & 0.18 & 0.071 & & 0.055 & 0.018 & & 0.018 & 0.0073 \\
        \hline
        \multicolumn{11}{|c|}{CMB $T + E$ priors (with and without post-Born corrections)} \\
        \hline
        && \multicolumn{3}{c|}{CMB lensing} & \multicolumn{3}{c|}{Gal. lensing} & \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
        \hline
        Par & prior & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
        \hline
$m_\nu$    & 0.22 & 0.074 & 0.12 & 0.073 & 0.14 & 0.10 & 0.086 & 0.043 & 0.090 & 0.037 \\
\rowcolor{blue!10}
           &      & & 0.13 & 0.073 & & 0.12 & 0.099 & & 0.10 & 0.038 \\
$w_0$      & 0.058 & 0.046 & 0.053 & 0.046 & 0.048 & 0.039 & 0.021 & 0.032 & 0.037 & 0.017 \\
\rowcolor{blue!10}
           &       & & 0.053 & 0.046 & & 0.042 & 0.022 &  & 0.039 & 0.018 \\
$\sigma_8$ & 0.015 & 0.014 & 0.015 & 0.013 & 0.013 & 0.011 & 0.0079 & 0.0097 & 0.010 & 0.0061 \\
\rowcolor{blue!10}
           &       & & 0.015 & 0.014 &  & 0.012 & 0.0077 & & 0.011 & 0.0063 \\
$\Omega_m$ & 0.015 & 0.014 & 0.015 & 0.014 & 0.013 & 0.011 & 0.0087 & 0.010 & 0.0096 & 0.0055 \\
\rowcolor{blue!10}
           &       & & 0.015 & 0.014 & & 0.011 & 0.0085 & & 0.010 & 0.0057 \\
        \hline
    \end{tabular}

\caption{Parameter constraints for different combinations of weak lensing information. $C$ stands for power spectra, and $B$ for bispectra. All surveys are assumed to be stage 4 surveys with multipole range $2 \leq \ell \leq 2000$. We use the weak and CMB priors described in table \ref{tab:fiducialpars}. The blue colored rows show constraints from bispectra with post-Born corrections. Post-Born corrections are not taken into account for the lensing power spectra as mentioned earlier, so these values are omitted in the blue rows.}
\label{tab:paramconstraintstight}
\end{table}

% CONFIDENCE ELLIPSES

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/param_constraints_tight_cmb_weak_prior.pdf}
    \caption{Parameter constraints and confidence ellipses for $\mu_\nu$, $w_0$, $\sigma_8$, and $\Omega_m$. Using ``stage 4'' noise with multipole range from $2$ to $2000$. The colored plots show CMB lensing power and/or bispectrum constraints. We use the weak priors listed in table \ref{tab:fiducialpars}. The confidence ellipses are for $1\sigma$ certainty. They show approximate degeneracies in the information gained from a survey if they are ``stretched''. When the information of ellipses with degeneracies in different directions is combined, the degeneracies are removed and the constraints typically become much better on the relevant parameters. The black plots show the constraints using all lensing information, i.e. CMB lensing spectra, galaxy lensing spectra, and all cross spectra.}
    \label{fig:paramconstraintstightcmb}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/param_constraints_tight_cmb_cmb_prior.pdf}
    \caption{Same as figure \ref{fig:paramconstraintstightcmb}, except we use the CMB temperature and polarization based priors listed in table \ref{tab:fiducialpars}.}
    \label{fig:paramconstraintstightcmbcmbprior}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/param_constraints_tight_gal_weak_prior.pdf}
    \caption{Same as figure \ref{fig:paramconstraintstightcmb}, except here the colored plots show galaxy lensing constraints.}
    \label{fig:paramconstraintstightgal}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/param_constraints_tight_gal_cmb_prior.pdf}
    \caption{Same as figure \ref{fig:paramconstraintstightcmbcmbprior}, except here the colored plots show galaxy lensing constraints.}
    \label{fig:paramconstraintstightgalcmbprior}
\end{figure}

% \begin{figure}
%     \centering
%     \tiny
%     \begin{tabular}{lrrrrr}
%         \hline
%          Par        &   CMB + Gal Powersp &   CMB + Gal Bisp &   CMB Power- + Bisp &   Gal Power- + Bisp &   CMB + Gal Power- + Bisp \\
%         \hline
%          $H_0$      &      102.8657993427 &     1.9383910910 &        9.6317154638 &        5.2219296015 &              0.9395018958 \\
%          $\sigma_8$ &       31.8199604416 &     3.8499102909 &        9.1631304725 &        4.5639390699 &              1.8357116464 \\
%          $\Omega_m$ &      172.9940493951 &     5.6526767874 &       27.2943604041 &       14.4462387878 &              3.5663315554 \\
%         \hline
%     \end{tabular}
% \caption{Same as table \ref{tab:paramconstraintstight}, except for a smaller selection of parameters well constrained by lensing.}
% \label{tab:paramconstraintstight}
% \end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[width=\textwidth]{figures/param_constraints_tight.pdf}
%     \caption{Same as \ref{fig:paramconstraintsall}, except showing constraints for a set of parameters that can be determined particularly accurately with weak lensing surveys. We chose Hubble's constant, $H_0$, baryonic + dark matter density, $\Omega_m$, and the root mean square density variation when the matter distribution is smoothed on a scale of 8 megaparsecs, commonly known as $\sigma_8$.}
%     \label{fig:paramconstraintstight}
% \end{figure}

\pagebreak

\section{Discussion and conclusion}\label{sec:discussion}
In this work, we have presented parameter forecasts from CMB lensing and galaxy lensing power and bispectra, considering experimental parameters representative of stage 3 and stage 4 CMB and galaxy surveys. Our analysis shows that while lensing power spectra are detectable at high significance, the inclusion of bispectra generally offers significant improvements when considering parameter constraints when considering an LCDM+ model of the universe. Notably, both CMB and galaxy bispectra are found to be detectable, even with stage 3 experimental noise levels.

The primary impact of incorporating bispectrum information is a tightening of cosmological parameter constraints, particularly for parameters such as $H_0$, $\sigma_8$, and $\Omega_m$, and in breaking key parameter degeneracies. We found that for sufficiently low noise levels, characteristic of stage 4 surveys, the bispectra themselves can offer constraining power comparable to that of the power spectra, due to nonlinear structure formation on small scales being better detected by these future surveys. Furthermore, the combination of CMB and galaxy lensing probes, especially when both power spectra and bispectra are utilized, further enhances constraints and aids in mitigating degeneracies. If all information were to be combined\footnote{And using the CMB $T+E$ prior.} we find particularly tight constraints on neutrino mass and dark energy equation of state in particular, i.e. $\sigma(m_\nu) = 55$ meV and $\sigma(w_0) = 0.017$. However, lensing power spectra alone, even for stage 4 surveys, appear insufficient to competitively constrain an 8-parameter $\Lambda\text{CDM} + w_0 + \sum m_\nu$ model without any priors.

There are a number of limitations to the results presented in this paper. The limitations inherent to the Fisher matrix formalism and to non-physical noise models such as those considered in this paper are already well known. The main limitations specific to this work are:
\begin{itemize}
     \item The fitting formula used for the nonlinear matter bispectrum (from \cite{bispfit}) has an estimated accuracy of only up to about 10 percent. This intrinsic inaccuracy in the model for $B^{\delta}$ will propagate to the lensing bispectra.

     \item Certain cosmological parameters that affect nonlinear large-scale structure, such as neutrino masses, likely alter the fitting parameters of the matter bispectrum formula when varied. In the literature, this is currently not considered, which can make the derivatives of the bispectrum with respect to these cosmological parameters less accurate.

     \item We assumed throughout this paper that lensing surveys would use data from multipoles as low as $\ell = 2$.  However, our analytical formulas to calculate lensing spectra utilize the Limber approximation, which is only robust for $\ell \gtrsim 50$. Avoiding the Limber approximation would complicate calculations, requiring knowledge of unequal-time matter power and bispectra\footnote{Technically, this would be computationally much more demanding due to the need to evaluate double integrals instead of single integrals}. Moving beyond the Limber approximation can be achieved using methods such as those explored in reference \cite{Chen:2021vba} and might offer a path to address this.

     \item As noted in appendix \ref{sec:fisher}, a Gaussian approximation was used for the covariance of the spectra estimators, i.e. we ignored any contributions not arising from standard Wick contractions. We know that the estimated lensing potential is not Gaussian in general, for two reasons: (i) the matter distribution is not Gaussian. (ii) Even with a Gaussian matter distribution, the quadratic estimator we used will not be Gaussian. Consider, for example, $\langle \hat \phi \hat \phi \hat \phi \rangle$. Each $\hat \phi$ is given as a sum over products of the lensing CMB anisotropies, $T$, $E$, $B$. Writing out each $\hat \phi$ in this way would yield terms such as $\langle TTTTTT\rangle$ (i.e. 6-point functions). These terms are non-zero. In fact, not only would they give products of 2-point functions due to Wick contractions, but also connected $n$-point functions with $n>2$ because the lensed fields are no longer Gaussian even if the original fields are. A next step could thus be to account for these additional contributions, perhaps through N-body and ray-tracing simulations. Notably, such simulations could potentially also allow for a more accurate calculation of the lensing spectra themselves, without relying on fitting formulas for the matter bispectrum or the Limber approximation, depending on simulation accuracy. Obviously, such simulations are expensive.

     \item Related to the previous point, for a completely correct analysis, we would also need to take into account correlations between a CMB prior and the lensing power and bispectra. Currently, we simply add Fisher matrices; however, taking these correlations into account means computing a covariance matrix for combinations of $T$, $E$, $\phi$ with nonzero entries for correlations between $T$/$E$ and $\phi$. For instance, consider the correlation between the estimator of the temperature power spectrum and the lensing power spectrum at arbitrary multipoles. We will get terms like $\langle TT\hat\phi\hat\phi \rangle$, which will be a sum over 6-point functions like $\langle TT TETE\rangle$ due to our quadratic estimator. These 6-point functions are clearly non-zero, as explained earlier.

     \item Another consideration is that non-Gaussianity of the true lensing field doesn't only change the covariance matrices used to calculate Fisher matrix elements, but also introduces an additional bias to the estimated lensing powerspectrum, denoted as $N^{(3/2)}_\ell$ \cite{B_hm_2016, B_hm_2018}. This is because the weights in quadratic estimator from ref. \cite{Hu_2002} are calculated while assuming a gaussian lensing field. The bias is significant at the percent level for CMB lensing, however it may be larger for cross correlations of the CMB lensing field with low redshift mass tracers (such as galaxy lensing). Fitting $N^{(3/2)}_\ell$ to survey data will loosen constraints for the cosmological parameters.

     \item Other than the lensing powerspectrum and bispectrum, there are also other statistics from the lensing field that can be used for parameter constraints. For example, ref. \cite{Liu_2016} showed that the lensing convergence one-point probability distribution function (PDF) and peak counts can tighten constraints in the $\Omega_m$ - $\sigma_8$ plane by $\sim30\%$ over just using the lensing powerspectrum for a stage 3 experiment\footnote{Specifically, for the AdvACT experiment.}. It is thus clear that future work could combine the power- and bispectrum with these additional statistics to further tighten forecasted constraints.

    \item The CMB lensing noise was estimated via a quadratic estimator. While standard, such estimators may not be strictly optimal, especially with complex foregrounds, non-Gaussianities and on small scales. Estimators that, given our stage 4 noise parameters, perform better do exist; for example, the iterative estimator developed in \cite{Smith_2012} can lead to a significantly better signal-to-noise ratio for the CMB lensing bispectrum, as shown in \cite{Namikawa_2016}. Using such an iterative estimator may also result in the CMB bispectrum being able to significantly improve parameter constraints when combined with lensing power spectra and a CMB $T+E$ prior.

    \item To calculate the Fisher matrices, derivatives of the lensing spectra with respect to cosmological parameters were computed using a central difference formula.  This introduces errors (estimated at most a few percent, though larger for neutrino mass derivatives, see appendix \ref{sec:derivatives}). Functions such as the matter power spectrum and fitting parameters for the matter bispectrum were approximated using pre-generated data and interpolation, also introducing errors of at most a few percent. Currently, these numerical errors are not considered significant because this paper primarily aims to provide a qualitative understanding of the utility of CMB and galaxy weak lensing spectra.  However, for precise quantitative results, these factors would need to be revisited and improved.

    \item We did not model foreground contamination such as the cosmic infrared background (CIB), thermal Sunyaev-Zel'dovich (tSZ) effect, or point sources. These foregrounds can bias lensing reconstructions and bispectra, particularly in CMB–galaxy cross-correlations, and should be accounted for in future analyses. Baryonic feedback was also not accounted for. Given that future surveys will probe the matter distribution on smaller scales, this will become increasingly relevant \cite{McCarthy_2022}, particularly for estimating neutrino mass \cite{PhysRevD.101.063534}. Bias due to baryonic feedback can be mitigated, e.g. by using the Principal Component Analysis method \cite{s2024impactcosmologydependencebaryonic}, discarding small scale $T$-modes, or fitting additional parameters that capture the effects of baryonic feedback \cite{McCarthy_2022}. Some of these methods may be able to mitigate baryonic feedback effects without increasing the uncertainty of parameter constraints \cite{McCarthy_2021}. Finally, there are still other well-known sources of uncertainty that we did not take into account, such as photometric redshift errors for galaxies or instrumental systematics.

\end{itemize}

Despite these limitations, this work provides valuable insight into the potential of future weak lensing surveys. Addressing these limitations will be crucial for obtaining robust cosmological constraints from upcoming observational data.

\section*{Acknowledgements}
We would like to thank Toshiya Namikawa for providing extra details on his paper on the CMB lensing bispectrum \cite{Namikawa_2016}, which, in particular, allowed us to cross-check (intermediate) results. We would also like to thank Antony Lewis for confirming the correctness of the formulas for post-Born corrections. The authors would like to thank Alex van Engelen for detailed comments and suggestions to improve the manuscript. Finally, the authors thank Luna Jansma, who wrote an MSc thesis on the same subject, which served as the foundation of this study.

\printbibliography

\appendix

\section{Weak Lensing}\label{sec:weaklensing}
\subsection{Perturbed Photon Paths}
We work in the conformal Newtonian gauge and with natural units. Denoting conformal time and conformal radial distance by $\eta$ and $\chi$, respectively, the perturbed line element in FLRW spacetime is given by
\begin{equation}
    \d s^2 = a^2(\eta)((1 + 2\Psi_N) \d \eta^2 - (1 + 2\Phi_N) \gamma_{ij}\d x^i \d x^j),
\end{equation}
where $\gamma_{ij}$ is the unperturbed line element
\begin{equation}
    \gamma_{ij} = \d x^i \d x^j = \d \chi^2 + f^2_K(\chi)(\d \theta^2 + \sin ^2 \theta \d \phi ^2),
\end{equation}
and $f_K(\chi)$ is the comoving angular diameter distance. We will hereafter only consider a flat universe so that $f_K(\chi) = \chi$. Weak lensing of a point source can be quantified by looking at the deflection field $\mathbf d (\hat{\mathbf n}) = \mathbf \theta_{\text{obs}} - \mathbf \theta_{\text{true}}$, i.e. the (small) difference between the angle at which we see the object and the angle at which we would see the object had no lensing occurred. To first order in $\Psi_N$ and $\Phi_N$, the deflection is given as \cite{dodelson2020modern}
\begin{equation}
    \mathbf d(\hat{\mathbf n}) = -2 \int_0^{\chi_*}\d \chi \frac{\chi_* - \chi}{\chi_*\chi}\nabla_{\hat{\mathbf n}}\Psi(\chi\hat{\mathbf n};\eta_0 - \chi),
\end{equation}
where $\Psi$ is the Weyl Potential, $\Psi := (\Psi_N - \Phi_N)/2$, and $\chi_*$ is the conformal distance to the source. $\nabla_{\hat{\mathbf n}}$ is the derivative along the axes orthogonal to the line of sight. The above equation can be written in terms of the lensing potential, $\psi$, as $\mathbf d (\hat{\mathbf n}) = \nabla_{\hat{\mathbf n}} \psi(\hat{\mathbf n})$, with
\begin{equation}
    \psi(\hat {\mathbf n}) := -2 \int_0^{\chi_*}\d \chi \frac{\chi_* - \chi}{\chi_*\chi}\Psi(\chi\hat{\mathbf n};\eta_0 - \chi).
\end{equation}

If the source is instead distributed over radial distance according to some distribution function $p(\chi)$, with $p(\chi)$ normalized to integrate to 1, the $(\chi_* - \chi)/(\chi_*\chi)$ factor is changed as
\begin{equation*}
    \frac{\chi_*-\chi}{\chi_*\chi} \rightarrow W(\chi) := \int_\chi^{\infty}\d\chi' p(\chi')\frac{\chi'-\chi}{\chi'\chi}.
\end{equation*}
$W(\chi)$ is then called the window function. In the most general case, the lensing potential is thus given by
\begin{equation}
    \psi(\hat {\mathbf n}) := -2 \int_0^{\infty}\d \chi W(\chi)\Psi(\chi\hat{\mathbf n};\eta_0 - \chi).
\end{equation}
The integration limit is sometimes also taken to be the surface of the last scattering, as any window function vanishes after that distance. In the case of CMB lensing we can take $p(\chi')=\delta(\chi'-\chi_*)$, in which case the window function reduces to $H(\chi_* - \chi)(\chi_*-\chi)/(\chi_*\chi)$, with $H(\chi)$ the Heaviside step function.

% From here of on work in first order in the scalar potentials $\Phi_N$ and $\Psi_N$. For lensing we consider null-geodesics so $\d s^2 = 0$ and we can rewrite the perturbed line element as
% \begin{equation}
%     \d \hat s^2 = (1 + 4\Psi)\d\eta^2 - \gamma_{ij} \d x^i \d x^j
% \end{equation}
% with $\Psi$ the \textbf{Weyl Potential} given by $\Psi := (\Psi_N - \Phi_N)/2$. A null geodesic $x^\mu(\hat\lambda)$ in terms of its affine parameter ($\hat \lambda$) satisfies the geodesic equation
% \begin{equation}
%     \frac{\d^2 x^\mu}{\d \hat\lambda^2} + \Gamma^\mu_{\nu\rho} \frac{\d x^\nu}{\d \hat\lambda}\frac{\d x^\rho}{\d\hat\lambda} = 0,
% \end{equation}
% with $\hat g_{\mu\nu}(\d x^\mu/\d \hat\lambda)(\d x^\nu / \d \hat \lambda) = 0$.

% In the unperturbed there are incoming radial solutions of the form $\chi = \eta_0 - \eta$. We can express $\hat\lambda$ in terms of $\eta$ using The 0-component of the geodesic equation,
% \begin{equation}
%     \frac{\d^2\eta}{\d\hat\lambda^2} + 2(\frac{\d\eta}{\d\hat\lambda})^2\frac{\d\Psi}{\d\eta}+2\frac{\d\eta}{\d\hat\lambda}\frac{\d x^i}{\d\hat\lambda}\frac{\partial\Psi}{\partial x^i} = 0,
% \end{equation}
% where the derivative $\d\Psi/\d\eta = \partial_\eta\Psi+(\d x^i/\d\eta)\partial_i\Psi$ is along the perturbed ray. This allows us to rewrite the other (spatial) geodesic equations as
% \begin{equation}
%     \frac{\d^2x^i}{\d\eta^2} - 2\frac{\d x^i}{\d\eta}(\frac{\d\Psi}{\d\eta} + \frac{\d x^j}{\d\eta}\frac{\partial\Psi}{\partial x^j})+2\gamma^{ij}\frac{\partial\Psi}{\partial x^j}+\bar\Gamma^i_{jk}\frac{\d x^j}{\d\eta}\frac{\d x^k}{\d \eta} = 0,
% \end{equation}
% where $\bar\Gamma^i_{jk}$ are the connection coefficients of the unperturbed three geometry $\gamma_{ij}$. Without loss of generality we consider an observer located at the origin of the spatial coordinates, meaning we are interested in rays that end at $x^i = 0$. In that case $\d \chi/\d \eta = -1 + O(\Psi)$, $\d \theta/\d\eta = O(\Psi)$ and $\d \phi/\d\eta = O(\Psi)$ (with $\theta$ and $\phi$ angular spatial coordinates). For the case of such rays we can evaluate the connection coefficients and rewrite the spatial geodesic equations as
% \begin{gather}
%     \frac{\partial^2\chi}{\d\eta^2}+2\frac{\d\Psi}{\d\eta}=0,\\
%     \frac{\d^2\theta}{\d\eta^2}-2\frac{\d\ln f_K(\chi)}{\d\chi}\frac{\d\theta}{\d\eta}+\frac{2}{f^2_K(\chi)}\frac{\partial\Psi}{\partial\theta}=0,\\
%     \frac{\d^2\phi}{\d\eta^2} - 2\frac{\d \ln f_K(\chi)}{\d\chi}\frac{\d\phi}{\d\eta}+\frac{2}{f^2_K(\chi)}\frac{1}{\sin^2(\theta)}\frac{\partial\Psi}{\partial\phi}=0.
% \end{gather}
% Integrating twice and using the null condition we can rewrite the the first equation as
% \begin{equation}
%     \chi = \eta_0 - \eta - 2\int_{\eta_0}^\eta\Psi\d\eta',
% \end{equation}
% where the integral is along the ray. Since we are working at first order in $\Psi$ we can evaluate the integral along the unperturbed path, this is known as the \textbf{Born approximation}.\textit{Next follows a short discussion of why perturbations in observation time can be ignored, include?} Doing the same for the other two equations gives
% % TODO: make n a proper vector (bold)
% \begin{gather}
%     \theta(\eta_0 - \chi_*)=\theta_0 - \int_0^{\chi_*}\d\chi\frac{f_K(\chi_* - \chi)}{f_K(\chi_*)f_K(\chi)}2\frac{\partial}{\partial \theta}\Psi(\chi\hat n, \eta_0 - \chi),\\
%     \phi(\eta_0 - \chi_*)=\theta_0 - \int_0^{\chi_*}\d\chi\frac{f_K(\chi_* - \chi)}{f_K(\chi_*)f_K(\chi)}\frac{2}{\sin^2\theta}\frac{\partial}{\partial \phi}\Psi(\chi\hat n, \eta_0 - \chi),
% \end{gather}
% where $\theta_0$ and $\phi_0$ label the line of sight $\hat n$. The displacement vector $\mathbf{\alpha}$ indicates the infinitesimal displacement of the lensed rays from their unpetrubed observed angles and is related to $\theta$ and $\phi$ through
% $$
% \alpha_\theta = \theta - \theta_0, \quad \alpha_\phi = \sin^2\theta(\phi - \phi_0).
% $$
% An object 
% \begin{equation}
%     \mathbf \alpha = -2 \int_0^{\chi_*}\d \chi \frac{f_K(\chi_* - \chi)}{f_K(\chi_*)f_K(\chi)}\nabla_{\hat{\mathbf n}}\Psi(\chi\hat{\mathbf n};\eta_0 - \chi).
% \end{equation}
% When working in the weak field limit (i.e. small spacetime curvature) the $00$ component of Einsteins equations give us a poisson equation \cite{dodelson2020modern},
% \begin{equation}
%     -\nabla^2\Psi(\textbf{x}, \eta) =  4\pi G a(\eta)^2\delta\rho(\mathbf x, \eta)
% \end{equation}
% where $K$ is $0$, $1$, or $-1$ depending on the curvature of the universe, $\Delta = \partial_x^2 + \partial_y^2 + \partial_z^2$ is the 3 dimensional laplace operator, $\delta\bar\rho$ is the comoving total energy perturbation (evaluated in the rest-frame of the total energy). %TODO: proper definition for anisotropic stress and explain why we can ignore it

\subsection{Convergence and Shear}
The magnification matrix is defined as
\begin{equation}
    A_{ij} := \delta_{ij} + \frac{\partial}{\partial n_j}d_i(\hat{\mathbf n}).
\end{equation}
This matrix can be decomposed in the following form, which immediately gives us definitions for the \textbf{convergence}, $\kappa$, \textbf{shear}, $\gamma_1$ and $\gamma_2$, and \textbf{rotation}, $\omega$:
\begin{equation}
    A_{ij}(\hat{\mathbf n}) = \begin{pmatrix}
        1 - \kappa - \gamma_1 & -\gamma_2 + \omega \\
        -\gamma_2 - \omega & 1 - \kappa + \gamma_1
    \end{pmatrix}.
\end{equation}
In the weak lensing regime and under the Born approximation, $A$ is a symmetric matrix by definition, and $\omega$ vanishes; we will ignore it from here on out. Intuitively, $A$ tells you how a small patch in the sky transforms due to lensing. If we change the unlensed direction of a light source by $\delta\hat{\mathbf n}$, then the corresponding change in direction in the lensed image can be calculated as
\begin{equation}
\hat{\mathbf{n}} + \delta \hat{\mathbf{n}} \rightarrow \hat{\mathbf{n}} + \delta \hat{\mathbf{n}} + \mathbf d (\hat{\mathbf{n}} + \delta \hat{\mathbf{n}}) = \hat{\mathbf{n}} + \mathbf d (\hat{\mathbf n}) + \delta \hat{\mathbf n} + A_{ij}(\delta \hat{\mathbf{n}})_j.
\end{equation}
For an image of the sky, $A_{ij}$ thus introduces distortion. Note that $|A_{ij}| = (1 - \kappa)^2 + \omega^2 - |\gamma|^2 = 1 - 2\kappa + O(\kappa^2, \gamma^2, \omega^2)$. We can thus interpret $\kappa$ as telling us about the overall magnification of the source. The $\gamma_i$ represents the area-preserving distortion, i.e. stretching and squeezing in a specific direction.

We can relate $\kappa$ and $\gamma$ directly to the lensing potential as
\begin{equation}
    \kappa = \frac{1}{2}\nabla^2\psi, \quad \gamma_1 = \frac{1}{2}(\partial_{n_1}^2 - \partial_{n_2}^2)\psi, \quad \gamma_2 = \partial_{n_1}\partial_{n_2} \psi.
\end{equation}

It is shown in appendix \ref{sec:shear} that
\begin{equation}
    \gamma := \gamma_1 + i\gamma_2 = \frac{1}{2}\eth_1(\eth_0\psi)
\end{equation}
where the spin raising operator, $\eth_s$ acts on a spin $s$ function defined on $S^2$ to create a spin $s+1$ function.
$\eth_s$ can be written in spherical coordinates $(\theta, \phi)$ as\footnote{We use the physics convention for the definition of $\theta$ and $\phi$ here.}
\begin{equation}
    \eth_s = -\sin^s\theta(\partial_\theta + \frac{i}{\sin\theta})\frac{1}{\sin^s\theta}.
\end{equation}
In this context, a spin $s$ function refers to a function $_sf(\theta, \phi)$ that transforms under any rotation of coordinates by picking up a phase factor $e^{is\alpha}$, with $\alpha$ the angle of the rotation, i.e.
\begin{equation}
    f'(\theta', \phi') = e^{is\alpha}f(\theta, \phi).
\end{equation}
Shear is thus a spin 2 function, which can be checked by noting that rotating a galaxy image stretched and squeezed through weak lensing by $180$ degrees gives the same stretching and squeezing, i.e. the same shear.

% The spin weighted spherical harmonics, $_sY_l^m$ are eigenfunctions of the raising/lowering operator, in particular
% \begin{equation}
%     \eth_sY_l^m = \sqrt{(l-s)(l+s+1)} _{s+1}Y_l^m.
% \end{equation}
The spherical harmonics are eigenfunctions of $\nabla^2$ and the spin raising and lowering operators. Using this property, the corresponding relations in spherical harmonic space can be shown to be 
\begin{equation*}
    \kappa_{lm} = \frac{l(l+1)}{2}\psi_{lm}, \quad \gamma_{lm} = \frac{\sqrt{(l-1)l(l+1)(l+2)}}{2} \psi_{lm}.
\end{equation*}

\section{Weak Lensing Statistics}\label{sec:weaklensstats}
\subsection{Lensing Potential Power spectrum}
The lensing potential can be decomposed into spherical harmonics as
\begin{equation}
    \psi(\hat{\mathbf n}) = \sum_{\ell m} \psi_{\ell m}Y_{\ell m}(\hat{\mathbf n}).
\end{equation}
On the other hand, consider the decomposition of $\Psi$ in Fourier modes with the Fourier convention $\Psi(\mathbf x, \eta)=\int\frac{\d^3 \mathbf k}{(2\pi)^{3}}\Psi(\mathbf k, \eta)e^{i\mathbf k \cdot\mathbf x}$,
\begin{equation}
    \psi(\hat{\mathbf n}) = -2 \int_0^{\chi_*}\d \chi W(\chi) \int\frac{\d^3 \mathbf k}{(2\pi)^{3}}\Psi(\mathbf k, \eta_0 - \chi)e^{i\mathbf k \cdot\hat{\mathbf n}\chi}.
\end{equation}
We can then relate the multipole modes of $\psi$ to the Fourier modes of $\Psi$ through % TODO: check sign convention here
\begin{gather}
    \psi_{lm} = \langle Y_\ell^m | \psi \rangle = \int\d^2\hat{\mathbf n} Y_\ell^m(\hat{\mathbf n})^* \psi(\hat{\mathbf n})\\ 
    = -2\int\d^2\hat{\mathbf n} Y_l^m(\hat{\mathbf n})^* \int_0^{\chi_*}\d \chi W(\chi) \int\frac{\d^3 \mathbf k}{(2\pi)^{3}}\Psi(\mathbf k, \eta_0 - \chi)e^{i\mathbf k \cdot\hat{\mathbf n}\chi}
\end{gather}
Now, define the power spectrum as
\begin{equation}
    \langle \Psi(\mathbf k, \eta)\Psi(\mathbf k',\eta')\rangle = \frac{2\pi^2}{k^3}P_\Psi(k, \eta, \eta')\delta(\mathbf k + \mathbf k'),
\end{equation}
with $\eta$ denoting the conformal time. This gives 
%TODO: define the window function
\begin{equation}
    \langle \psi(\hat{\mathbf n})\psi(\hat{\mathbf n}') \rangle = 4\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi')\int \frac{\d^3\mathbf k}{(2\pi)^6}\frac{2\pi^2}{k^3}P_\psi(k,\eta_0 - \chi, \eta_0-\chi')e^{i\mathbf k\cdot\hat {\mathbf n}\chi}e^{-i\mathbf k\cdot\hat {\mathbf n}'\chi'},
\end{equation}
where we used that $\eta = \eta_0 - \chi$ along the unperturbed photon path (this is known as the Born approximation), with $\eta_0$ the time at which the light ray hits Earth. 
We can use the result
\begin{equation}
    e^{i\mathbf k \cdot \hat {\mathbf n}\chi}=4\pi\sum_{\ell m}i^\ell j_\ell(k\chi )Y_\ell^m(\hat{\mathbf n})^*Y_\ell^m(\hat{\mathbf k}) = 4\pi\sum_{\ell m}i^\ell j_\ell(k\chi )Y_\ell^m(\hat{\mathbf n})Y_\ell^m(\hat{\mathbf k})^*, \label{eq:complexexp}
\end{equation}
where $j_l$ is the spherical Bessel function, to rewrite the above equation. Using both versions of the identity above, we immediately get a factor $Y_\ell^m(\hat{\mathbf k})Y_{\ell'}^{m'}(\hat{\mathbf k})^*$ in our integral. We can factor the differential element of $\d^3\mathbf k$ into a radial and angular part as $k^2\d k\d^2\Omega_k$, with $\Omega_k$ the solid angle, to apply the orthonormality condition of the spherical harmonics. Note that we take the same sequence of steps a number of times in other parts of the derivations of the lensing spectra. We thus obtain
\begin{gather}
    \langle \psi(\hat{\mathbf n})\psi(\hat{\mathbf n}') \rangle = 4(4\pi)^2\sum_{ll'mm'}i^{l-l'}\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi')\\\times\int \frac{k^2\d k}{(2\pi)^6}\frac{2\pi^2}{k^3}j_\ell(k\chi)j_{\ell'}(k\chi')P_\psi(k,\eta_0 - \chi, \eta_0-\chi')Y_{\ell m}(\hat{\mathbf n})Y_{\ell'm'}(\hat{\mathbf n}')^*\delta_{\ell \ell'}\delta_{mm'}. \label{eq:lenspotcorrexpanded}
\end{gather}
The angular power spectrum is defined similarly to the power spectrum, i.e.
\begin{equation}
    \langle \psi_{\ell m}\psi_{\ell'm'}^* \rangle = \delta_{\ell \ell'}\delta_{mm'}C_\ell^\psi.
\end{equation}
Note that the correlation is independent of $m$ and $m'$. %TODO: explain how this relates to statistical isotropy
We can thus read from equation \ref{eq:lenspotcorrexpanded} that
\begin{gather}
    C_l^\psi = 4(4\pi)^2\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi')\int \frac{k^2\d k}{(2\pi)^6}\frac{2\pi^2}{k^3}j_\ell(k\chi)j_{\ell}(k\chi')P_\psi(k,\eta_0 - \chi, \eta_0-\chi'),
\end{gather}
which can be simplified to
\begin{gather}
    C_\ell^\psi = \frac{2}{\pi^2}\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi')\int k^2\d kj_\ell(k\chi)j_{\ell}(k\chi')\frac{P_\psi(k,\eta_0 - \chi, \eta_0-\chi')}{k^3}.
\end{gather}

To further evaluate the integral we will make the Limber approximation. The Bessel functions peak sharply at $l=x$\footnote{Some sources use $x\approx l+1/2$ instead, which is slightly more accurate for larger scales (low $l$) and slightly less accurate for smaller scales.}, with the peak being increasingly sharp for higher $l$. Similarly to $\delta(x-x_0)f(x)=\delta(x-x_0)f(x_0)$, we thus take $j_l(k\chi)f(k)\approx j_l(k\chi)f(l/\chi)$. The Bessel functions satisfy an orthogonality condition,
\begin{equation}
    \int k^2\d k j_\ell(k\chi)j_\ell(k\chi') = \frac{\pi}{2\chi^2}\delta(\chi-\chi').
\end{equation}
In combination with the Limber approximation, we thus find
\begin{equation}
    \int k^2\d k j_\ell(k\chi)j_\ell(k\chi')f(k) \approx \frac{\pi}{2\chi^2}\delta(\chi-\chi')f(\ell/\chi).
\end{equation}
%TODO: how high? quantify limber approx, perhaps using convergence x LSS paper
It allows us to write the Limber-approximate angular spectrum as
\begin{gather}
    C^\psi_{\ell} = \frac{2}{\pi^2}\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi') \frac{\pi}{2\chi^2}\delta(\chi-\chi')\frac{\chi^3}{\ell^3}P_\psi(\ell/\chi,\eta_0 - \chi, \eta_0-\chi')\\
    = \frac{1}{\ell^3\pi}\int_0^{\chi_*}\chi\d \chi W(\chi)^2 P_\psi(\ell/\chi,\eta_0 - \chi, \eta_0-\chi). \label{eq:lpsintermsofpsi}
\end{gather}
% The power spectrum of the potential can be related to the power spectrum of the matter density using the Poisson equation %TODO: create ref
% For a flat universe in atter or dark energy domination we thus find the relation
% \begin{equation}
%     P_\Psi(k;\eta, \eta)=\frac{9\omega_m^2(\eta)H^4(\eta)}{8\pi^2} \frac{P_m(k, \eta)}{k},
% \end{equation}
% where we use the shorthand $P_m(k, \eta):=P_m(k,\eta,\eta)$. The lensing power spectrum is thus finally given in terms of the matter power spectrum as
% \begin{equation}
%     C_l^\psi=\underbrace{9\omega_m^2(\eta)H^4(\eta)}_{:=C(\eta)}\frac{1}{l^4}\int_0^{\chi_*}\d\chi\chi^2 W(\chi)^2P_m(l/\chi, \eta_0-\chi).
% \end{equation}
% We will henceforth denote the $\eta$ dependent factor outside of the integral as $C(\eta)$.
\subsection{Lensing potential bispectrum}
The derivation of the bispectrum proceeds similarly to that of the power spectrum. We aim to compute the bispectrum of the lensing potential fields of 3 (possibly distinct sources), $\psi_1$, $\psi_2$, $\psi_3$. 
\begin{gather*}
    \langle (\psi_1)_{\ell_1m_1}(\psi_2)_{\ell_2m_2}(\psi_3)_{\ell_3m_3} \rangle = 
    \prod_i\br{-2\int\d^2\hat{\mathbf n_i}(Y^{m_i}_{\ell_i}(\hat{\mathbf n_i}))^*\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}e^{i\mathbf k_i \cdot\hat{\mathbf n_i}\chi_i}} \\
    \times \langle \prod_i\Psi(\mathbf k_i, \eta_0 - \chi_i) \rangle.
\end{gather*}
Defining the bispectrum of the gravitational potential as
\begin{equation*}
    \langle \prod_{i=1,2,3}\Psi(\mathbf k_i, \eta_0 - \chi_i) \rangle = (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}),
\end{equation*}
we rewrite the lensing potential bispectrum as
\begin{gather*}
    \langle (\psi_1)_{\ell_1m_1}(\psi_2)_{\ell_2m_2}(\psi_3)_{\ell_3m_3} \rangle = 
    \prod_i\br{-2\int\d^2\hat{\mathbf n_i}(Y^{m_i}_{\ell_i}(\hat{\mathbf n_i}))^*\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}e^{i\mathbf k_i \cdot\hat{\mathbf n_i}\chi_i}}\\ \times (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}).
\end{gather*}
Now using equation \eqref{eq:complexexp} to rewrite the complex exponential:
%TODO: check \ell m here
\begin{gather*}
    \langle (\psi_1)_{\ell_1m_1}(\psi_2)_{\ell_2m_2}(\psi_3)_{\ell_3m_3} \rangle \\ = 
    \prod_i\br{-2\int\d^2\hat{\mathbf n_i}(Y^{m_i}_{\ell_i}(\hat{\mathbf n_i}))^*\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}4\pi\sum_{\ell m}i^{\ell}j_{l}(k_i\chi_i)Y_{\ell}^{m}(\hat{\mathbf n_i})Y_{\ell}^{m}(\hat{\mathbf k_i})^*}\\
    \times (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}) \\
    = \sbr{\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}4\pi i^{\ell_i}j_{\ell_i}(k_i\chi_i)Y_{\ell_i}^{m_i}(\hat{\mathbf k_i})^*}} (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}).
\end{gather*}
We can write the 3D Dirac delta function in terms of spherical harmonics as
\begin{equation*}
    \delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3) = 8\int\d^3\mathbf x\prod_{i=1,2,3}\br{ \sum_{\ell_jm_j} i^{\ell_j}j_{\ell_j}(k_ix)Y_{\ell_j}^{m_j}(\hat{\mathbf k_i}) Y_{\ell_j}^{m_j}(\hat{\mathbf x})^* }.
\end{equation*}
This results in
\begin{gather*}
    \langle (\psi_1)_{\ell_1m_1}(\psi_2)_{\ell_2m_2}(\psi_3)_{\ell_3m_3} \rangle= \prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}4\pi i^{\ell_i}j_{\ell_i}(k_i\chi_i)Y_{\ell_i}^{m_i}(\hat{\mathbf k_i})^*} \\
    \times (2\pi)^38\int\d^3\mathbf x\prod_{i}\br{ \sum_{\ell m} i^{\ell}j_{\ell}(k_ix)Y_{\ell}^{m}(\hat{\mathbf k_i}) Y_{\ell}^{m}(\hat{\mathbf x})^* }B^\Psi(\{k_i\}, \{\eta_0-\chi_i\})\\
    = (2\pi)^38\int\d^3\mathbf x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{k_i^2\d k_i}{(2\pi)^{3}}4\pi (-1)^{\ell_i}j_{\ell_i}(k_i\chi_i)j_{\ell_i}(k_ix) Y_{\ell_i}^{m_i}(\hat{\mathbf x})^*} B^\Psi(\{k_i\}, \{\eta_0-\chi_i\})
\end{gather*}
The angular part of the $\mathbf x$ integral can be evaluated using the identity
\begin{gather*}
    \int \mathrm{d}\Omega_{\hat{n}} \, Y_{\ell_1 m_1}^*(\hat{x}) Y_{\ell_2 m_2}^*(\hat{n}) Y_{\ell_3 m_3}^*(\hat{n}) = (-1)^{m_1 + m_2 + m_3} 
    \int \mathrm{d}\Omega_{\hat{n}} \, Y_{\ell_1 -m_1}(\hat{n}) Y_{\ell_2 -m_2}(\hat{n})Y_{\ell_3 -m_3}(\hat{n}) \\
    = (-1)^{m_1 + m_2 + m_3} 
    \sqrt{\frac{(2\ell_1 + 1)(2\ell_2 + 1)(2\ell_3 + 1)}{4\pi}} \begin{pmatrix} \ell_1 & \ell_2 & \ell_3 \\ 0 & 0 & 0 \end{pmatrix}
    \begin{pmatrix} \ell_1 & \ell_2 & \ell_3 \\ -m_1 & -m_2 & -m_3 \end{pmatrix} \equiv A_{\mathbf{l}}^{\mathbf{m}},
\end{gather*}
giving
\begin{gather*}
    \langle (\psi_1)_{\ell_1m_1}(\psi_2)_{\ell_2m_2}(\psi_3)_{\ell_3m_3} \rangle 
    = (2\pi)^3 8 A_{\mathbf l}^{\mathbf m}\int x^2\d x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{k_i^2\d k_i}{(2\pi)^{3}}4\pi (-1)^{\ell_i}j_{\ell_i}(k_i\chi_i)j_{\ell_i}(k_ix)} \\ \times B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}).
\end{gather*}
Now applying the Limber approximation again:
\begin{gather*}
    \langle (\psi_1)_{\ell_1m_1}(\psi_2)_{\ell_2m_2}(\psi_3)_{\ell_3m_3} \rangle 
    = (2\pi)^3 8 A_{\mathbf l}^{\mathbf m}\int x^2\d x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\frac{1}{(2\pi)^{3}}\frac{\pi}{2\chi_i^2}\delta(x-\chi_i)4\pi (-1)^{\ell_i}} \\ \times B^\Psi(\{\ell_i/\chi_i\}, \{\eta_0-\chi_i\}) \\
    = (2\pi)^3 8 A_{\mathbf l}^{\mathbf m}\int \chi^2\d \chi\prod_i\br{-2 W_i(\chi)\frac{1}{(2\pi)^{3}}\frac{\pi}{2\chi^2}4\pi (-1)^{\ell_i}} B^\Psi(\{\ell_i/\chi\}, \eta_0-\chi).
\end{gather*}
Finally, we aim to rewrite the above in terms of the angular bispectrum of the lensing potential.

The definition for the bispectrum of any set of randomly distributed spherical harmonic components $X^k_{\ell m}$ is \cite{Hu2000}
\begin{equation*}
    \langle (X_1)_{\ell_1m_1}(X_2)_{\ell_2m_2}(X_3)_{x_3m_3} \rangle = \begin{pmatrix}
        \ell_1 & \ell_2 & \ell_3 \\ m_1 & m_2 & m_3
    \end{pmatrix}
    B_{\ell_1\ell_2\ell_3}^{X_1X_2X_3}.
\end{equation*}
Note the independence on $m_i$, this necessarily follows from statistical isotropy. % TODO: add a short explanation here
If $m_1+m_2+m_3\neq 0$, the associated Wigner-3j symbol vanishes and the bispectrum is set to zero.
Also note that in this definition we have immediately generalized to include cross correlation between different fields $X_1$, $X_2$, $X_3$. This is relevant when we look at cross-correlations between CMB and galaxy lensing.
6758
Using the above definition and the symmetry property
\begin{equation*}
    \begin{pmatrix}
        \ell_1 & \ell_2 & \ell_3\\
        -m_1 & -m_2 & -m_3
      \end{pmatrix}
      =
      (-1)^{\ell_1+\ell_2+\ell_3}
      \begin{pmatrix}
        \ell_1 & \ell_2 & \ell_3\\
        m_1 & m_2 & m_3
      \end{pmatrix},
\end{equation*}
we find
\begin{gather*}
    B^{\psi_1\psi_2\psi_3}_{\ell_1\ell_2\ell_3}
    = (-1)^{\ell_1+\ell_2+\ell_3}
    \sqrt{\frac{(2\ell_1 + 1)(2\ell_2 + 1)(2\ell_3 + 1)}{4\pi}} \begin{pmatrix} \ell_1 & \ell_2 & \ell_3 \\ 0 & 0 & 0 \end{pmatrix}
    (2\pi)^3 8 \\
    \times \int \chi^2\d \chi\prod_i\br{-2 W_i(\chi, \chi_*)\frac{1}{(2\pi)^{3}}\frac{\pi}{2\chi^2}4\pi (-1)^{\ell_i}} B^\Psi(\{\ell_i/\chi\}, \{\eta_0-\chi\}),
\end{gather*}
where we were able to drop the $(-1)^{m_1+m_2+m_3}$ factor due to the bispectrum vanishing if that sum does not equal $0$, as mentioned earlier. When all $m_i$ equal zero, the Wigner 3j-symbol gains a number of useful properties % TODO: discuss all of them
In particular, it vanishes if $\ell_1+\ell_2+\ell_3$ is odd, meaning we can drop the $(-1)^{\ell_1+\ell_2+\ell_3}$ factor. Additionally, cancelling common factors then gives
\begin{align}
    B^{\psi_1\psi_2\psi_3}_{\ell_1\ell_2\ell_3}
    =
    -\sqrt{\frac{(2\ell_1 + 1)(2\ell_2 + 1)(2\ell_3 + 1)}{4\pi}} \begin{pmatrix} \ell_1 & \ell_2 & \ell_3 \\ 0 & 0 & 0 \end{pmatrix} 8 \int \frac{\d \chi }{\chi^4} W_1(\chi)W_2(\chi)W_3(\chi) B^\Psi(\{l_i/\chi\}, \eta_0-\chi). \label{eq:lbsintermsofpsi}
\end{align}

% \begin{gather*}
%     = \prod_i\br{-2\int\d^2\hat{\mathbf n_i}(Y^{m_i}_{l_i}(\hat{\mathbf n_i}))^*\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3/2}}4\pi\sum_{l_i'm_i'}i^{l_i'}j_{l_i'}(k_i\chi_i)Y_{l_i'}^{m_i'}(\hat{\mathbf n_i})Y_{l_i'}^{m_i'}(\hat{\mathbf k_i})^*}\\ 
%     \times (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3) B_\Psi(k_1, k_2, k_3) \\
%     = \prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3/2}}4\pi i^{l_i}j_{l_i}(k_i\chi_i)Y_{l_i}^{m_i}(\hat{\mathbf k_i})^*}\\ 
%     \times (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3) B_\Psi(k_1, k_2, k_3) \\
%     = \prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3/2}}4\pi i^{l_i}j_{l_i}(k_i\chi_i)Y_{l_i}^{m_i}(\hat{\mathbf k_i})^*}\\ 
%     \times (2\pi)^3\sbr{ \int \d^3 \mathbf{x} \prod_i2\sum_{l_i'm_i'}i^{l_i'}j_{l_i'}(k_ix) Y_{l_i'm_i'}(\hat{\mathbf k}_i) Y_{l_i'm_i'}(\hat{\mathbf x})^* } B_\Psi(k_1, k_2, k_3) \\
%     = \prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{k_i^2\d k_i}{(2\pi)^{3/2}}4\pi i^{l_i}j_{l_i}(k_i\chi_i)}\\ 
%     \times (2\pi)^3\sbr{ \int \d^3 \mathbf{x} \prod_i2i^{l_i}j_{l_i}(k_ix) Y_{l_im_i}(\hat{\mathbf x})^* } B_\Psi(k_1, k_2, k_3) \\
%     = A_{\mathbf l}^{\mathbf m}\int x^2\d x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{k_i^2\d k_i}{(2\pi)^{3/2}}4\pi i^{l_i}(2i^{l_i})j_{l_i}(k_i\chi_i)j_{l_i}(k_ix)}
%     (2\pi)^3  B_\Psi(k_1, k_2, k_3) \\
%     = A_{\mathbf l}^{\mathbf m}\int x^2\d x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\frac{1}{(2\pi)^{3/2}}4\pi i^{l_i}(2i^{l_i})\frac{\pi}{2\chi_i^2}\delta(x-\chi_i)}
%     (2\pi)^3  B_\Psi(\ell_1/\chi, \ell_2/\chi, \ell_3/\chi) \\
%     = A_{\mathbf l}^{\mathbf m}\int_0^{\chi_*} x^2\d x\prod_i\br{-2 W(x, \chi_*)\frac{1}{(2\pi)^{3/2}}4\pi i^{l_i}(2i^{l_i})\frac{\pi}{2x^2}}
%     (2\pi)^3  B_\Psi(\ell_1/\chi, \ell_2/\chi, \ell_3/\chi) \\
%     = A_{\mathbf{l}}^{\mathbf{m}} (-1)^{1 + \ell_1 + \ell_2 + \ell_3} 8 (2\pi)^{9/2} \int_0^{\chi_*} \frac{[W(\chi, \chi_*)]^3}{\chi^4} \, B_\Psi\left( \frac{\ell_1}{\chi}, \frac{\ell_2}{\chi}, \frac{\ell_3}{\chi} \right ) \, \mathrm{d}\chi.
% \end{gather*}

% TODO: mention that the definition of the bispectrum is different form power spectrum of Psi in terms of normalization constant
% TODO: explain 3d dirac delta in terms of spherical harmonics (eq 78 thesis Luna)
% TODO: explain that we used that int d^2 Omega Y Y Y = A_vec l^vec m (eq 82 thesis Luna)
% using that d^3 k = k^2 d k d^2 Omega_k
% TODO: fix this
\subsection{Gravitational potential spectra in terms of matter spectra}
% TODO: fix references and citations 
We can rewrite equations \ref{eq:lpsintermsofpsi} and \ref{eq:lbsintermsofpsi} in terms of the matter spectra instead of the $\psi$ spectra using the Poisson equation. This allows us to numerically evaluate these lensing spectra using CAMB. The density contrast is defined as
\begin{equation}
    \delta(\mathbf x) := \frac{\rho(\mathbf x)-\bar \rho}{\bar\rho},
\end{equation}
and the matter spectra are defined in terms of the Fourier transformed density contrast $\delta(\mathbf k)$ as 
\begin{gather*}
    \langle \delta(\mathbf k, \eta) \delta(\mathbf k', \eta)^* \rangle = (2\pi)^3\delta(\mathbf k - \mathbf k')P^\delta(\mathbf k, \eta),\\
    \langle \delta(\mathbf k_1, \eta) \delta(\mathbf k_2, \eta) \delta(\mathbf k_3, \eta) \rangle = (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\delta(k_1, k_2, k_3, \eta).
\end{gather*}
The mean matter density of the universe, \( \bar{\rho} \) is given as 
$$
\bar{\rho}(\eta) = \frac{3 \Omega_m H_0^2}{8 \pi G}\frac{1}{a(\eta)^3},
$$
where $a(\eta)$ is the only time-dependent factor on the right-hand side.
The Poisson equation relates $\Psi$ to the density contrast as \cite{dodelson2020modern}
% in modern cosm eq 6.41 they say - nabla^2 Phi = ..., however here we use Psi instead and Psi = - Phi in our regime so hence why we don't take the - sign
\begin{equation}
    \nabla^2\Psi(\mathbf x) = 4\pi G a^2\br{\frac{3\Omega_m H_0^2}{8\pi G}\frac{1}{a^3}}\delta(\mathbf x) = \frac{3\Omega_m H_0^2}{2} \frac{1}{a}\delta(\mathbf x) \implies \Psi(k, \eta) = -\frac{3\Omega_m H_0^2}{2}\frac{1}{a}\frac{\delta(k, \eta)}{k^2},
\end{equation}
where $\Psi(k,\eta)$ and $\delta(k,\eta)$ are functions in Fourier space. For the power- and bispectra, we find
\begin{gather*}
    \langle \Psi(\mathbf k, \eta)\Psi^*(\mathbf k',\eta)\rangle = \frac{2\pi^2}{k^3}C^\Psi(k, \eta)\delta(\mathbf k - \mathbf k') \implies C^\Psi(k,\eta) = \frac{1}{k}(9\Omega_m^2H_0^4\pi) \frac{1}{a^2}C^\delta(k,\eta),\\
    \langle \Psi(\mathbf k_1, \eta)\Psi(\mathbf k_2,\eta)\Psi(\mathbf k_3,\eta)\rangle = -(2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(k_1,k_2,k_3, \eta)\\ 
    \implies B^\Psi(k_1,k_2,k_3, \eta) = -\frac{1}{k_1^2k_2^2k_3^2}\br{\frac{3\Omega_m H_0^2}{2}}^3 \frac{1}{a^3}B^\delta(\{k_i\}, \eta).
\end{gather*}
Finally, we obtain:
\begin{empheq}[box=\fbox]{align*}
    C^{\psi_X\psi_Y}_\ell
    & = \frac{9}{\ell^4}\Omega_m^2H_0^4\int_0^{\chi_*} \chi^2 \d\chi a(\eta_0-\chi)^{-2}W_X(\chi)W_Y(\chi)P^\delta(\ell/\chi,\eta_0-\chi),\\
    B^{\psi_X\psi_Y\psi_Z}_{\ell_1\ell_2\ell_3} &= \sqrt{\frac{(2\ell_1 + 1)(2\ell_2 + 1)(2\ell_3 + 1)}{4\pi}} \begin{pmatrix} \ell_1 & \ell_2 & \ell_3 \\ 0 & 0 & 0 \end{pmatrix} \frac{27}{\ell_1^2\ell_2^2\ell_3^2}\Omega_m^3H_0^6\\
    & \quad \times \int \chi^2\d \chi a(\eta_0-\chi)^{-3}W_X(\chi)W_Y(\chi)W_Z(\chi)  B^\delta(\{\ell_i/\chi\}, \eta_0-\chi).
\end{empheq}

\section{Fisher Matrix Analysis} \label{sec:fisher}
\subsection{Determining uncertainty in experimental parameters}

The Fisher matrix formalism is used to find a lower bound on the constraints we can place on experimental parameters. It combines the Cramer-Rao Inequality \cite{casella2002statistical} with the assumption that we have unbiased estimators following a Gaussian distribution \cite{dodelson2020modern}. In particular, denoting the Fisher matrix by $F_{\theta_i\theta_j}$, the parameters as $\theta_i$, and their estimators as $\hat\theta_i$, it can be shown that
\begin{equation}
    \text{Var}(\hat\theta_i) \geq \br{F^{-1}}_{\theta_i\theta_i}.
\end{equation}
In the case of $n$ measurements whose outcomes are realizations of random variables $x_i$, each with associated mean $\mu_{x_i}$, the Fisher matrix is given as
\begin{equation}
    F_{\theta_i\theta_j} := \sum_{p, q=1}^n \frac{\partial\mu_{x_p}}{\partial\theta_i}(\tilde\theta_k)(\text{Cov}^{-1})_{x_px_q}(\tilde \theta_k)\frac{\partial\mu_{x_q}}{\partial\theta_j}(\tilde\theta_k),
\end{equation}
where Cov is the covariance matrix associated with the random vector $(x_1, ..., x_n)$, $\text{Cov}_{x_px_q}:=\text{Cov}(x_p,x_q)$. The derivative of the mean measurement outcomes and measurement covariances will in general, depend on the true, unknown values of the experimental parameters. We therefore evaluate these quantities for our best guess of the experimental parameters given some outside information. These are known as the ``fiducial'' values.

%%%%%%%%%%%%%%%%%%%%%

\subsection{Fisher matrices for power- and bispectra with multiple tracers}
For power spectra, the definition of the Fisher matrix gives 
\begin{equation*}
    F_{\alpha\beta} = \sum_{l_{min}\leq \ell, \ell' \leq \ell_{max}}\sum_{[XY][X'Y']} \partial_\alpha C_{\ell}^{XY} \br{\Cov^{-1}}^{XY, X'Y'}_{\ell, \ell'} \partial_\beta C^{X'Y'}_{\ell'}.
\end{equation*}
Here, the covariance matrix is given as
\begin{equation*}
    \Cov^{XY, X'Y'}_{\ell, \ell'} = \langle \hat C^{XY}_{\ell} \hat C^{X'Y'}_{\ell'} \rangle.
\end{equation*}
The estimator of the power- or bi-spectrum is the product of the estimators of the appropriate tracers, e.g. $\hat C^{XY}_{\ell} = \hat X(\ell) \hat Y(\ell)$. The sum over $[XY]$ and $[X'Y']$ denotes a sum over possible tracer combinations \textit{without} counting permutations of tracer configuration. This is because permutations are not distinct signals, i.e. $\hat X(\ell) \hat Y(\ell) = \hat Y(\ell) \hat X(\ell)$. In fact, if we were to count these permutations as distinct signals, we would get identical columns and/or rows in our covariance matrix, making inversion impossible:
\begin{equation*}
    \langle \hat X(\ell) \hat Y (\ell) \hat X'(\ell') \hat Y' (\ell') \rangle = \langle \hat Y (\ell) \hat X (\ell) \hat X'(\ell') \hat Y' (\ell') \rangle, \quad \forall X', Y', \ell'. 
\end{equation*}

To evaluate the covariance matrix, we again assume that the estimators are Gaussian distributed so that we can apply a Wick contraction, as is commonly done \cite{tegmark1997cosmic}. In this case we get
\begin{equation*}
    \Cov^{XY, X'Y'}_{\ell, \ell'} = \frac{1}{2\ell+1}\delta_{\ell \ell'}\br{\tilde C^{XX'}_\ell \tilde C^{YY'}_\ell + \tilde C^{XY'}_\ell \tilde C^{YX'}_\ell}.
\end{equation*}
Two remarks are in order.
\begin{enumerate}
    \item By definition
    \begin{equation*}
        \langle X_{\ell m}X'_{\ell'm'} \rangle = (2\pi)^3 \delta_{\ell\ell'}\delta_{mm'}C_\ell^{XX'},
    \end{equation*}
    so you can average over the measurements done for different values of $m$, i.e. $X_{\ell(-\ell)}$, $X_{\ell(-\ell+1)}$, $\cdots$, $X_{\ell(\ell-1)}$, $X_{\ell \ell}$. This results in the $(2\ell+1)^{-1}$ factor in the power spectrum covariance. 
    \item The tilde is used to denote the power spectrum as calculated earlier, plus the noise power spectrum, $N^{XX'}_\ell$, which is the power spectrum of the noise associated with the estimator of the field. This is where experimental noise is incorporated into the calculation. 
\end{enumerate}
Under the Gaussian approximation, the covariance matrix vanishes except for $3\times 3$ block matrices (in the case of 2 tracers) on the diagonal. The Fisher matrix is then
\begin{equation*}
    F_{\alpha\beta} = \sum_{\ell}\sum_{[XY][X'Y']} \partial_\alpha C_{\ell}^{XY} \br{\Cov^{-1}_\ell}^{XY, X'Y'} \partial_\beta C^{X'Y'}_{\ell}.
\end{equation*} 
$\Cov^{-1}_\ell$ here denotes the inverse of the block matrix at $l$.

Next, we consider the Fisher matrix for bispectra measurements.
\begin{equation*}
    F_{\alpha\beta} = \sum_{\text{distinct signals}}\sum_{\text{distinct signals prime}} B^{XYZ}_{\ell_1\ell_2\ell_3}\br{\Cov^{-1}}^{XYZ,X'Z'Y'}_{\ell_1\ell_2\ell_3,\ell_1'\ell_2'\ell_3'}B^{X'Y'Z'}_{\ell_1'\ell_2'\ell_3'}.
\end{equation*} 
Counting only distinct signals requires more care compared to the power spectra. The rule is that $B^{XYZ}_{\ell_1\ell_2\ell_3}$ and $B^{X'Y'Z'}_{\ell_1'\ell_2'\ell_3'}$ are not distinct signals if there exists a permutation $\sigma$ that simultaneously maps $X'Y'Z'$ to $XYZ$ and $\ell_1'\ell_2'\ell_3'$ to $\ell_1\ell_2\ell_3$. It turns out that we can write a sum over distinct signals explicitly as
\begin{gather*}
    \sum_{\text{distinct signals}} = \underbrace{\sum_{\ell_1=\ell_2=\ell_3}\sum_{[XYZ]}}_{\text{sum }1} + \underbrace{\sum_{\ell_1=\ell_2\neq \ell_3}\sum_{[XY]Z}}_{\text{sum }2} + \underbrace{\sum_{\ell_1<\ell_2<\ell_3}\sum_{XYZ}}_{\text{sum }3}.
\end{gather*}
With the same definition again for the $[\cdot]$ notation. For example:
\begin{gather*}
    % \{\{XYZ\}|X,Y,Z\in\{\psi_\kappa, \psi_\gamma\}\} = \{\psi_\kappa\psi_\kappa\psi_\kappa, \psi_\kappa\psi_\kappa\psi_\gamma, \psi_\kappa\psi_\gamma\psi_\gamma, \psi_\gamma\psi_\gamma\psi_\gamma\}, \\
    \{[XY]Z|X,Y,Z\in\{\psi_\kappa, \psi_\gamma\}\} = \{\psi_\kappa\psi_\kappa\psi_\kappa, \psi_\kappa\psi_\gamma\psi_\kappa, \psi_\gamma\psi_\gamma\psi_\kappa,\psi_\kappa\psi_\kappa\psi_\gamma, \psi_\kappa\psi_\gamma\psi_\gamma, \psi_\gamma\psi_\gamma\psi_\gamma\}.
\end{gather*}

It follows to show that the sets that these sums sum over form a partition of the set of all distinct signals. Clearly, all 3 sets are pairwise disjoint (no common elements) because of the criteria for the $l_i$'s. To show that their union covers the set of distinct signals, start by considering an arbitrary signal. Its $l$ configuration will trivially correspond to exactly one of the three sums. If it is sum 1, then we are free to permute the $XYZ$'s by virtue of the $l$'s being identical, so we will be able to match the $XYZ$ configuration to one of the elements of $\{\{XYZ\}\}$. Similarly, if the $l$ configuration corresponds to sum 2, then we are free to permute the $XY$ configuration to match with one of the elements of $\{\{XY\}Z\}$. The $Z$ value does not matter because any $Z$ value is accounted for. For sum 3, we can argue that we can switch around the order of the $l$'s to satisfy $\ell_1 < l2 < l3$ and the corresponding $XYZ$ configuration will be accounted for in sum 3 because all $XYZ$ combinations are counted. Finally, it is simple to verify that no distinct signal is counted more than once within each sum.

To calculate the elements of the covariance matrix, consider the following. Every element of the Fisher matrix can be seen as an inner product weighted by the inverse covariance matrix. We can choose how we order the vectors\footnote{The entries are the derivatives of the bispectra}. We organize the vectors according to the sum of 1, 2, and 3 parts first. Then by $l$ configuration. Within each $l$ configuration, we can choose any ordering for the $XYZ$ configurations. The covariance matrix now becomes a block matrix with each block corresponding to an $l_i$ and $l_i'$ configuration. When wick contracting using the Gaussian approximation, every block matrix where $(\ell_1,\ell_2,\ell_3)\neq (\ell_1',\ell_2',\ell_3')$ vanishes. It can then be shown that the entries of each block matrix are given as
\begin{gather*}
    \br{\Cov_{\ell_1\ell_2\ell_3}}^{XYZ,X'Y'Z'} = \tilde C^{XX'}_{\ell_1}\tilde C^{YY'}_{\ell_2}\tilde C^{ZZ'}_{\ell_3} + \delta_{\ell_1\ell_2}\tilde C^{XY'}_{\ell_1}\tilde C^{YX'}_{\ell_2}\tilde C^{ZZ'}_{\ell_3} +\delta_{\ell_2\ell_3}\tilde C^{XX'}_{\ell_1}\tilde C^{YZ'}_{\ell_2}\tilde C^{ZX'}_{\ell_3} \\ + \delta_{\ell_3\ell_1}\tilde C^{XZ'}_{\ell_1}\tilde C^{YY'}_{\ell_2}\tilde C^{ZX'}_{\ell_3} +\delta_{\ell_1\ell_2}\delta_{\ell_2\ell_3}\br{\tilde C^{XY'}_{\ell_1}\tilde C^{YZ'}_{\ell_2}\tilde C^{ZX'}_{\ell_3} +\tilde C^{XZ'}_{\ell_1}\tilde C^{YX'}_{\ell_2}\tilde C^{ZY'}_{\ell_3}}.
\end{gather*}
With our ordering, this means that the covariance matrix is again a diagonal block matrix, now with blocks of size $4 \times 4$ (sum 1), $6 \times 6$ (sum 2), and $8 \times 8$ (sum 3).

\subsection{Explicit form for inverse covariance matrix}
The Fisher matrix above can be significantly simplified and written as
\begin{equation*}
F_{\alpha\beta} = \sum_{\ell_1 \leq \ell_2 \leq \ell_3} \frac{\mathcal P _{\ell_1\ell_2\ell_3}}{6}
\sum_{XYZ}\sum_{X'Y'Z'} 
\partial_\alpha B^{X Y Z}_{\ell_1 \ell_2 \ell_3} 
(\tilde C^{-1})^{X X'}_{\ell_1}
(\tilde C^{-1})^{Y Y'}_{\ell_2}
(\tilde C^{-1})^{Z Z'}_{\ell_3}
\partial_\beta B^{X' Y' Z'}_{\ell_1 \ell_2 \ell_3}
\end{equation*}
where, in the case of two tracers, 
\begin{equation*}
    C_{l} := \begin{pmatrix}
        \tilde C^{\psi_1\psi_1}_l & \tilde C^{\psi_1\psi_2}_l \\
        \tilde C^{\psi_1\psi_2}_l & \tilde C^{\psi_2\psi_2}_l
    \end{pmatrix}
\end{equation*}
and $\mathcal P_{\ell_1\ell_2\ell_3}$ is defined as the number of distinct permutations that can be made with $\ell_1\ell_2\ell_3$. This form was, for example, used in \cite{Kalaja_2021}\footnote{Note that in \cite{Kalaja_2021} this is based on a previous equation summing over \textit{all} $l_i$ (so including permutations of each configuration) which is missing a factor of $1/6$.}.

To show that the above is the same as the formula for the Fisher matrix given earlier, first partition the sum in the same way and collect all terms that fit in the different categories.
\begin{align*}
    F_{\alpha\beta} &= \sum_{\ell_1=\ell_2=\ell_3} \sum_{[XYZ]}\sum_{[X'Y'Z']} \partial_{\alpha} B^{XYZ}_{\ell_1 \ell_2 \ell_3}\br{\frac{\mathcal P _{\ell_1\ell_2\ell_3}}{6}\sum_{d. p. XYZ} \sum_{d. p. X'Y'Z'}(\tilde C^{-1})^{X X'}_{\ell_1}
    (\tilde C^{-1})^{Y Y'}_{\ell_2}
    (\tilde C^{-1})^{Z Z'}_{\ell_3}}
    \partial_\beta B^{X' Y' Z'}_{\ell_1 \ell_2 \ell_3} \\
    &+ \sum_{\ell_1=\ell_2\neq \ell_3} \sum_{[XY]Z}\sum_{[X'Y']Z'} \partial_{\alpha} B^{XYZ}_{\ell_1 \ell_2 \ell_3}\br{\frac{\mathcal P _{\ell_1\ell_2\ell_3}}{6}\sum_{d. p. XY} \sum_{d. p. X'Y'}(\tilde C^{-1})^{X X'}_{\ell_1}
    (\tilde C^{-1})^{Y Y'}_{\ell_2}
    (\tilde C^{-1})^{Z Z'}_{\ell_3}}
    \partial_\beta B^{X' Y' Z'}_{\ell_1 \ell_2 \ell_3} \\
    &+ \sum_{\ell_1<\ell_2<\ell_3} \sum_{XYZ}\sum_{X'Y'Z'} \partial_{\alpha} B^{XYZ}_{\ell_1 \ell_2 \ell_3}(\tilde C^{-1})^{X X'}_{\ell_1}
    (\tilde C^{-1})^{Y Y'}_{\ell_2}
    (\tilde C^{-1})^{Z Z'}_{\ell_3}
    \partial_\beta B^{X' Y' Z'}_{\ell_1 \ell_2 \ell_3},
\end{align*}
where ``$d. p.$'' stands for ``distinct permutations''.
The entries above are then the entries of the inverses of the block matrices mentioned earlier. This can be checked. For example, for the $\ell_1 = \ell_2 = \ell_3$ sum the multiplication of block matrices corresponding to the same $l_i$ configuration can be written as:
\begin{gather*}
    \sum_{[X'Y'Z']} \br{\frac{\mathcal P _{\ell_1\ell_2\ell_3}}{6} \sum_{d. p. XYZ} \sum_{d. p. X'Y'Z'}(\tilde C^{-1})^{X X'}_{l}
    (\tilde C^{-1})^{Y Y'}_{l}
    (\tilde C^{-1})^{Z Z'}_{l}} \br{\tilde C^{X'X''}_l \tilde C^{Y'Y''}_l \tilde C^{Z'Z''}_l + \text{perms } X''Y''Z''} \\
    = \frac{\mathcal P _{\ell_1\ell_2\ell_3}}{6}\sbr{\br{\sum_{d. p. XYZ} \sum_{X'Y'Z'}(\tilde C^{-1})^{X X'}_{l}
    (\tilde C^{-1})^{Y Y'}_{l}
    (\tilde C^{-1})^{Z Z'}_{l}} \tilde C^{X'X''}_l \tilde C^{Y'Y''}_l \tilde C^{Z'Z''}_l} + \text{perms } X''Y''Z'' \\
    = \frac{\mathcal P _{\ell_1\ell_2\ell_3}}{6} \sum_{d. p. XYZ} \delta_{XX''}\delta_{YY''}\delta_{ZZ''} + \text{perms } X''Y''Z''
    = \frac{\mathcal P _{\ell_1\ell_2\ell_3}}{6} \delta_{[XYZ], [X''Y''Z'']} + \text{perms } X''Y''Z''\\
    = \delta_{[XYZ], [X''Y''Z'']} .
\end{gather*}
The sum over the different wick contractions will similarly cancel with the $\mathcal P_{\ell_1\ell_2\ell_3} / 6$ factor for the $\ell_1 = \ell_2 \neq \ell_3$ sum. For the $\ell_1 < \ell_2 < \ell_3$ sum, the proof is similar as well, except no cancellation is required.

The same type of simplification can be made in the Fisher matrix for the power spectrum, though it doesn't offer any significant benefits compared to our current $3\times 3$ block matrix form.

\subsection{Signal to Noise Ratio (SNR)}
To quantify the detectability of the lensing spectra, we introduce an overall amplitude of our signal, $A$, with fiducial value 1 as an experimental parameter and compute $F_{AA}$. Obviously, $\partial_A\br{A B^{XYZ}_{\ell_1\ell_2\ell_3}}|_{A=1} = B^{XYZ}_{\ell_1\ell_2\ell_3}$, so we find
\begin{equation*}
    \br{\frac{S}{N}}^2 := F_{AA} =
\sum_{XYZ, X'Y'Z'} \sum_{\ell_1\leq \ell_2 \leq \ell_3} \frac{\mathcal P _{\ell_1\ell_2\ell_3}}{6}
B^{X Y Z}_{\ell_1 \ell_2 \ell_3} 
(\tilde C^{-1})^{X X'}_{\ell_1}
(\tilde C^{-1})^{Y Y'}_{\ell_2}
(\tilde C^{-1})^{Z Z'}_{\ell_3}
B^{X' Y' Z'}_{\ell_1 \ell_2 \ell_3}.
\end{equation*}
The equation for the SNR of the power spectra is identical in form.

\subsection{Fisher matrix of power- + bispectra}
To compute the Fisher matrix of an experiment measuring both lensing power and bispectra, we are required to compute and invert the full covariance matrix. If we keep assuming that the measurements are close enough to Gaussian to be able to use Wick contractions to a good approximation, the full covariance matrix simplifies trivially. The correlation between a power and bispectrum estimator contains an odd (5) amount of fields and thus always vanishes. We are thus allowed to simply add the Fisher matrices of the power and bispectra to compute the combined Fisher matrix.

% \subsection{Fisher Matrices and Eigenvalues}
% As explained earlier, inverting the Fisher matrix gives us a covariance matrix of the estimators, $\hat \theta_i$ of our experimental parameters. For any vector $\mathbf v$,
% \begin{equation*}
%     \mathbf{v}^T_j \Cov_{ij} \mathbf v_i = \langle (v_i \hat \theta_i)^2 \rangle.
% \end{equation*}
% In particular, if the covariance matrix thus has some large eigenvalue $\lambda_L$ (large compared to the overal accuracy of our experiment), with eigenvector $v_i$, then the corresponding estimator, $v_i\hat \theta_i$, has a large variance and thus represents an approximate degeneracy in our parameters. Clearly, when looking at the Fisher matrix we thus instead look for the smallest eigenvalue.

% If we have such an abnormally small eigenvalue we are likely to run into problems with calculating and inverting the Fisher matrix due to numerical errors. These problems are that inversion becomes unstable and that the matrix as a whole might not be PSD because the small eigenvalue was (wrongly) calculated to be negative. These problems are likely because (1) the smallest eigenvalue is sensitive to numerical errors due to being small and (2) the errors on the eigenvalues is proportial to the condition number of the matrix \cite{horn2012matrix}, where the condition number is the ratio between the largest and smallest eigenvalue of the matrix \cite{trefethen1997numerical}. Specifically, for a matrix $A$ with some perturbation $\Delta A$, the corresponding  perturbation of eigenvalue $\lambda_i$ is bounded above as
% \begin{equation*}
%     |\Delta \lambda_i | \leq ||\Delta A || \kappa(A)|,
% \end{equation*}
% with $\kappa(A)$ the condition number.

% Before inverting a Fisher matrix, it is thus recommended to remove rows and columns corresponding to parameters that are poorly constrained by the experiment, which can be determined by checking for small eigenvalues and their eigenvectors.

% Computing the inverse is then reduced to computing the inverse of the individual block matrices. The block matrices for sum 1 and 2 don't have a particularly nice form (cite paper about multiple tracer fisher matrices for bispectra) but for sum 3 the inverse of the block matrix corresponding to $\ell_1\ell_2\ell_3$ is
% \begin{equation*}
%     (\Cov_{\ell_1\ell_2\ell_3}^{-1})^{XYZ, X'Y'Z'} = \hat C_{\ell_1}^{X,X'}\hat C_{\ell_2}^{Y,Y'}\hat C_{\ell_3}^{Z,Z'}.
% \end{equation*}
% \begin{equation*}
%     (\Cov)^{XYZ, X'Y'Z'} = C_{\ell_1}^{X,X'} C_{\ell_2}^{Y,Y'} C_{\ell_3}^{Z,Z'}.
% \end{equation*}
% As can be seen from
% \begin{equation*}
%     \sum_{X'Y'Z'} = \hat C_{\ell_1}^{X,X'}\hat C_{\ell_2}^{Y,Y'}\hat C_{\ell_3}^{Z,Z'} C_{\ell_1}^{X'X''}C_{\ell_2}^{Y'Y''}C_{\ell_3}^{Z'Z''} = \delta_{XX''}\delta_{YY''}\delta_{ZZ''} = \delta_{XYZ, X''Y''Z''}
% \end{equation*}

% It's worth mentioning that due to all $l$'s being distinct in each configuration of sum 3 the Wick contraction corresponding to each element of the covariance matrix can only be done in one way and we thus only have one term. In the sum 2 we get two terms due to $\ell_1$ and $\ell_2$ being identical. In sum 3 we similarly get 6 terms corresponding to all possible Wick contractions being nonzero.

% \subsection{Shear noise}
% In the absence of lensing, galaxies will generally still look elliptical rather than perfectly spherical. This \textbf{intrinsic ellipticity} is a major cause of noise when measuring lensing shear. We can define the ellipticity of a galaxy image (or any type of object, for that matter) in terms of \textbf{quadrupole moments} of the surface brightness as
% \begin{equation*}
% 	e_1 = \frac{I_{xx} - I_{yy}}{I_{xx} + I_{yy}}, \quad e_2 = \frac{2I_{xy}}{I_{xx} + I_{yy}},
% \end{equation*}
% where
% \begin{equation*}
% 	I_{ij} = \int \d x\d y I(x,y)ij, \quad i,j\in\{x,y\},
% \end{equation*}
% and the galaxy is centered on $(x,y)=0$.

% The connection with lensing shear is then made by taylor expanding and using the lensing matrix. If we have unlensed ellipiticities $e_i^\text{true}$ and lensed ellipticities $e_i^{\text{obs}}$, then \cite{dodelson_modern_cosmology}
% \begin{gather*}
% 	e_1^{\text{obs}} = \br{1-2\sbr{e_1^{\text{true}}\gamma_1+e_2^{\text{true}}\gamma_2}}e_1^{\text{true}} + 2\gamma_1,\\
% 	e_2^{\text{obs}} = \br{1-2\sbr{e_1^{\text{true}}\gamma_1+e_2^{\text{true}}\gamma_2}}e_2^{\text{true}} + 2\gamma_2.
% \end{gather*}
% In the case of weak lensing, $\gamma_i, \kappa \ll 1$, thus the above reduces to $\gamma_i \approx (e_i^{\text{obs}} - e^{\text{true}}_i)/2$. The analysis thusfar was quietly done in real space, but taking the harmonic transform of $\gamma_i$ is trivial under the approximation we just made. We thus find the error due to intrinsic ellipticity in harmonic space to equal $ \langle \gamma_{i,l}^2\rangle \approx \langle \br{e_{i,l}^\text{true}}^2\rangle/4$. In a real survey we don't measure the shear for individual galaxies, but instead average it over small patches of the sky. The size of each patch is chosen to balance between signal to noise ratio of the shear measurement and resolution of the shear map. If there are $N$ galaxies observed per patch, then the noise is given by
% $$
% \langle \br{\hat \gamma_{i,l}  - \gamma_{i,l}^{\text{true}}}^2\rangle \approx \frac{\langle \br{e_{i,l}^\text{true}}^2\rangle}{4N}.
% $$
% We don't know the exact value for the right hand side either, so we use the standard estimator for that. There is no correlation between two patches, i.e. the noise is white noise and thus scale independent. Finally, we want to look at the noise in $\gamma_l\gamma_{l'}^*$. We thus get
% \begin{equation*}
% 	C^{\gamma\gamma,\;\text{obs}}_l = C^{\gamma\gamma, \;\text{true}}_l + N_l^{\gamma\gamma},
% \end{equation*} 
% with
% \begin{equation*}
% 	 N_l^{\gamma\gamma} = \frac{1}{4}\frac{e_{\text{RMS}}}{N}, \quad e_{\text{RMS}}:=\sqrt{\frac{1}{N}\sum^N_{i=1}\sbr{(e_{1, i} - \bar e_1)^2 + (e_{2,i}-\bar e_2)^2}}.
% \end{equation*}

% Typically $e_{RMS} / 4 \sim 0.3$ \cite{Kilbinger_2015}.

% The EUCLID team aims for a noise of about 20 percent and thus require at least $\sim 4 \times 10^8$ galaxies measured per patch, giving an absolute error of \cite{2024}
% $$
% N_l^{\gamma\gamma} \approx 8 \times 10^{-10} \quad \text{(EUCLID)}.
% $$

% \subsection{Convergence Noise}
% \subsubsection{Optimal Estimator}
% In practice, the lensing potential cannot be directly measured but is instead estimated based on the temperature, $\Theta$, $E$-mode polarization, $E$, and $B$-mode polarization, $B$, fields.

% Define some axes $x$ and $y$ on a small patch of the sky and let $\mathbf E$ be the electric field of the CMB radiation, then the \textbf{linear polarization Stokes parameters} are defined as
% \begin{equation*}
% 	Q := E_xE^*_x - E_yE_y^*, \quad U := E_xE_y^* + E_yE_x^*.
% \end{equation*} 
% By looking at a sufficiently small patch of the sky we can apply the \textbf{flat sky limit}, in which case we take a 2D Fourier transform over the approximately flat patch instead of working with spherical harmonics. In this case we find $Q(\mathbf k)$ and $U(\mathbf k)$. In Fourier space we can instead work with the more convenient $E$ and $B$ modes, defined as
% \begin{gather*}
% 	E(\mathbf k)=\cos 2\phi_{\mathbf k} Q(\mathbf k) + \sin 2\phi_{\mathbf k} U(\mathbf k),\\
% 	B(\mathbf k)=-\sin 2\phi_{\mathbf k} Q(\mathbf k) + \cos 2\phi_{\mathbf k} U(\mathbf k),
% \end{gather*}
% where $\phi_\textbf{k}$ is the angle of $\mathbf k$ with the $(1,0)$  vector (i.e. $k_{\hat{\mathbf{x}}}$).
% $E$ corresponds to the curl free component of the polarization field and behaves like a scalar under transformations. $B$ corresponds to the rest of the polarization field and behaves like a semi-scalar under transformations. The Fourier components are thus related to the real space fields as
% \begin{gather*}
% \Theta(\mathbf{\hat{n}}) = \int \frac{d^2 k}{(2\pi)^2} \Theta(\mathbf{k}) e^{i \mathbf{k} \cdot \mathbf{\hat{n}}}, \\
% [Q \pm iU](\mathbf{\hat{n}}) = -\int \frac{d^2 k}{(2\pi)^2} \left[ E(\mathbf{k}) \pm iB(\mathbf{k}) \right] e^{\pm 2i \varphi_{\mathbf{k}}} e^{i \mathbf{k} \cdot \mathbf{\hat{n}}}, \\
% \phi(\mathbf{\hat{n}}) = \int \frac{d^2 K}{(2\pi)^2} \phi(\mathbf{K}) e^{i \mathbf{K} \cdot \mathbf{\hat{n}}}.
% \end{gather*}

% As discussed earlier, lensing acts in real space as 
% \begin{equation*}
%     X(\hat n) = \tilde X(\hat {\mathbf n} + \nabla \phi(\hat{\mathbf n})) \approx \tilde X(\hat{\mathbf n}) + \nabla \phi \cdot \nabla \tilde X |_{\hat{\mathbf{n}}},
% \end{equation*}
% with $X$ any of $\Theta$, $Q \pm iU$ and $\tilde X$ the corresponding \textit{unlensed} field. In Fourier space this naturally takes the form of a convolution. The changes in Fourier moments due to lensing are thus given by \cite{Hu_2002}
% \begin{gather*}
%     \delta \Theta(\mathbf{k}) = \int \frac{d^2 k'}{(2\pi)^2} \tilde{\Theta}(\mathbf{k'}) W(\mathbf{k'}, \mathbf{K}), \\[8pt]
%     \delta E(\mathbf{k}) = \int \frac{d^2 k'}{(2\pi)^2} 
%     \left[ \tilde{E}(\mathbf{k'}) \cos 2\varphi_{k'k} - \tilde{B}(\mathbf{k'}) \sin 2\varphi_{k'k} \right] W(\mathbf{k'}, \mathbf{K}), \\[8pt]
%     \delta B(\mathbf{k}) = \int \frac{d^2 k'}{(2\pi)^2} 
%     \left[ \tilde{B}(\mathbf{k'}) \cos 2\varphi_{k'k} + \tilde{E}(\mathbf{k'}) \sin 2\varphi_{k'k} \right] W(\mathbf{k'}, \mathbf{K}),
% \end{gather*}

% \noindent where \( \varphi_{k'k} \equiv \varphi_{k'} - \varphi_k \), \( \mathbf{K} = \mathbf{k} - \mathbf{k'} \), and
% \begin{gather*}
%     W(\mathbf{k}, \mathbf{K}) = -[\mathbf{k} \cdot \mathbf{K}] \phi(\mathbf{K}).
% \end{gather*}

% For power spectra in 2D Fourier space we use the definition
% \begin{equation*}
%     \langle x(\mathbf k)x'(\mathbf k') \rangle = (2\pi)^2\delta(\mathbf k + \mathbf k')C_k^{xx'}
% \end{equation*}
% unless stated otherwise. We will denote unlensed powerspectra with a tilde on top. $C^{xx'}(k)$ denotes the observed power spectrum is related to the true (but still lensed) spectrum as
% \begin{equation*}
%     C^{xx', \text{ obs}}(k) = B(k)^2C^{xx', \text{ true}}(k) + \delta_{xx'}N^x(k).
% \end{equation*}
% We are then able to express the power spectra of the lensed fields in terms of the unlensed fields as
% \begin{equation*}
%     \langle x(\mathbf k)x'(\mathbf k') \rangle = f_{xx'}(\mathbf k, \mathbf k') \phi(\mathbf k + \mathbf k'),
% \end{equation*}
% where $f_{xx'}$ is a function of 2 wavenumbers and unlensed power spectra. Specific values of $f$ for different combinations of $x$ and $x'$ can be found in ... %TODO: fix this
% We can then set up an estimator for $\phi$ using a weighted integral of the Fourier modes measured for $x$ and $x'$ by allowing for a filter function $F_{xx'}(\mathbf k, \mathbf k')$. This gives the estimator
% \begin{equation*}
%     \hat\phi_{xx'}(\mathbf k) = \frac{\int \frac{\d^2k'}{(2\pi)^2}x(\mathbf k')x'(\mathbf k - \mathbf k') F_{xx'}(\mathbf k', \mathbf k - \mathbf k')}{\int \frac{\d^2k'}{(2\pi)^2}f_{xx'}(\mathbf k', \mathbf k - \mathbf k') F_{xx'}(\mathbf k', \mathbf k - \mathbf k')}.
% \end{equation*}
% It's easily checked that $\langle \hat\phi_{xx'} \rangle = \phi$ regardless of the filter function used. Minimizing for the variance of the estimator gives
% \begin{equation}
%     F_{xx'}(\mathbf{k}_1, \mathbf{k}_2) = 
%     \frac{C_{k_1}^{x'x'} C_{k_2}^{xx} f_{xx'}(\mathbf{k}_1, \mathbf{k}_2) 
%     - C_{k_1}^{xx'} C_{k_2}^{xx'} f_{xx'}(\mathbf{k}_2, \mathbf{k}_1)}
%     {C_{k_1}^{xx} C_{k_2}^{x'x'} C_{k_1}^{x'x'} C_{k_2}^{xx} 
%     - \left( C_{k_1}^{xx'} C_{k_2}^{xx'} \right)^2}.
% \end{equation}
% With this estimator we assume that the unlensed powerspectra have been determined through other means (e.g. a simulation).
% With optimal weighting the noise becomes
% \begin{equation*}
%     \langle \hat\phi_{xy}(\mathbf K) \hat\phi_{x'y'}(\mathbf K') \rangle = (2\pi)^2\delta(\mathbf K - \mathbf K') [C^{\phi \phi}_K + K^{-2}N_{xyx'y'}(K)],
% \end{equation*}
% where
% \begin{equation*}
%     N_{xyx'y'}(K):=K^{-2}A_{xy}(K)A_{x'y'}(K)\int\frac{\d^2 k'}{(2\pi)^2}F_{xy}(\mathbf k', \mathbf k - \mathbf k')\br{
%         F_{x'y'}(\mathbf k', \mathbf k - \mathbf k') C_{k'}^{xx'}C^{yy'}_{k - k'} + F_{x'y'}(\mathbf k - \mathbf k', \mathbf k') C_{k'}^{xy'}C^{yx'}_{k - k'}
%     }.
% \end{equation*}

% \subsubsection{Noise}
% We partition the sky into a finite number of pixels of size (in solid angle) $\Omega_{\text{pix}}$. We take into account detector beam width and homogenous noise, but not boundary effects or inhomogeneous noise. %TODO: justify this or take it into account
% In real space the observed field relates to the true (but still lensed) field as
% \begin{equation*}
%     X^{\text{obs}}(\hat{\mathbf n}) = \int \d\Omega' X^{\text{true}}(\hat{\mathbf n}') B(\hat{\mathbf n}, \hat{\mathbf n}') + \eta(\hat{\mathbf n}),
% \end{equation*}
% where $B$ is the beam pattern and $\eta$ is the noise map stemming from any number of experimental sources of error.

% In the case of an isotropic uniform beam, the beam pattern can be written as a function of $\hat{\mathbf n} - \hat{\mathbf n}'$ and the Fourier transform gives 
% \begin{equation*}
%     X^{\text{obs}}(\mathbf k) = X^{\text{true}}(\mathbf k)B(\mathbf k) + \eta(\mathbf k),
% \end{equation*}
% with
% \begin{equation*}
%     B(\mathbf k) = e^{-\frac{k^2\theta_{\text{FWHM}}^2}{4\ln 2}},
% \end{equation*}
% and $\theta_{\text{FWHM}}$ the angle for full width at half minimum.
% By using a likelihood approach we obtain the estimator of the lensed power spectrum given by
% \begin{equation*}
%     \hat C^{XX}(k) = B(\mathbf k)^{-2}\br{X^{\text{obs}}(\mathbf k)X^{\text{obs}}(\mathbf k)^* - N(k)}
% \end{equation*}
% with variance
% \begin{equation*}
%     \Cov(\hat C^{XX}(k), \hat C^{XX}(k')) = 2\sbr{C(k) + N(k)B(k)^{-2}}\delta(\mathbf k - \mathbf k').
% \end{equation*}
% The noise as calculated in eq ... % TODO: ref 
% is thus given by using
% \begin{equation*}
%     C^{XY, \text{ obs}}(k) = \langle \hat C^{XY}(k) \rangle = B(\mathbf k)^{-2}(C)
% \end{equation*}
% \subsection{Power and bispectrum estimators}
% \subsection{Combining experimental and theoretical (shot noise?) uncertainty}
% %Let's say you are looking at a field $X_l$ and have some experimental error $E_l$, uncorrelated with $X_l$ and with mean $0$ (so no systematic error). Then
% %\begin{equation*}
% %	\langle (X+E)(X+E)^*\rangle = \langle X X^*\rangle + \langle EE^* \rangle.
% %\end{equation*}

% In the case of an error expressed as
% \begin{gather*}
% 	\text{``We measure the amplitude of the CMB lensing} \\\text{
% 	power spectrum at state-of-the-art 2.3\% precision, [...]"}
% \end{gather*}
% we can quantify this as
% \begin{equation*}
% 	\hat P^X_l = P_l^X(1+E),
% \end{equation*}
% where $E\sim \mathcal N(0, \sigma^2)$ and in the example above $\sigma = 0.023$. It thus follows that

%TODO: mention SNR

\section{Shear equals twice spin raised lensing potential}\label{sec:shear}

Consider a point on $S^2$, $(r_0, \theta_0, \phi_0)$, at which we observe some cosmological object. We can then define a set of cartesian coordinates $(\tilde r, \tilde y, \tilde x)$ as shown in figure \ref{fig:tildecoords}.

\begin{figure}[h!]
    \centering
    \input{figures/lensing_coords.tikz}
    \caption{
        $(\tilde r, \tilde y, \tilde x)$ coordinates defined for a point on the unit sphere. These act as ordinary cartesian coordinates but rotated such that, at the associated point on $S^2$, $\hat{\tilde r}$ points straight out of the unit sphere, $\hat{\tilde y}$ is parallel to the great arc with constant $\phi$ and $\hat{\tilde x}$ is parallel to the great arc with constant $\theta$. These coordinates are used to define the shear and convergence in terms of the lensing potential.
        }
    \label{fig:tildecoords}
\end{figure}

Note that it is not obvious whether to define these coordinates at the point where the lensed light hits $S^2$ or the unlensed light hits $S^2$. We will assume that lensing effects are sufficiently weak that either definition works. We can then express the tilde coordinates in terms of spherical coordinates either by applying a rotation matrix or by calculating the $r, \theta, \phi$ derivatives of $(x,y,z)$ coordinates at $(r_0, \theta_0, \phi_0)$ to find $\hat{\tilde{\mathbf r}}$, $\hat{\tilde{\boldsymbol{\theta}}}$, and $\hat{\tilde{\boldsymbol{\phi}}}$ and then take inner products. Regardless, we find
\begin{align}
    \tilde r &= r\sin\theta\cos\phi\sin\theta_0\cos\phi_0+r\sin\theta\sin\phi\sin\theta_0\sin\phi_0+r\cos\theta\cos\theta_0,\\
    \tilde y &= r\sin\theta\cos\phi\cos\theta_0\cos\phi_0+r\sin\theta\sin\phi\cos\theta_0\sin\phi_0-r\cos\theta\sin\theta_0,\\
    \tilde x &= -r\sin\theta\cos\phi\sin\phi_0+r\sin\theta\sin\phi\cos\phi_0.
\end{align}
This gives derivatives
\begin{align*}
    \frac{\partial}{\partial \theta} = & \left( r \cos \theta \cos \phi \sin \theta_0 \cos \phi_0 + r \cos \theta \sin \phi \sin \theta_0 \sin \phi_0 - r \sin \theta \cos \theta_0 \right) \frac{\partial}{\partial \tilde{r}} \\
    & + \left( r \cos \theta \cos \phi \cos \theta_0 \cos \phi_0 + r \cos \theta \sin \phi \cos \theta_0 \sin \phi_0 + r \sin \theta \sin \theta_0 \right) \frac{\partial}{\partial \tilde{y}} \\
    & + \left( -r \cos \theta \cos \phi \sin \phi_0 + r \cos \theta \sin \phi \cos \phi_0 \right) \frac{\partial}{\partial \tilde{x}}.\\
    \frac{\partial}{\partial \phi} = & \left( -r \sin \theta \sin \phi \sin \theta_0 \cos \phi_0 + r \sin \theta \cos \phi \sin \theta_0 \sin \phi_0 \right) \frac{\partial}{\partial \tilde{r}} \\
& + \left( -r \sin \theta \sin \phi \cos \theta_0 \cos \phi_0 + r \sin \theta \cos \phi \cos \theta_0 \sin \phi_0 \right) \frac{\partial}{\partial \tilde{y}} \\
& + \left( r \sin \theta \sin \phi \sin \phi_0 + r \sin \theta \cos \phi \cos \phi_0 \right) \frac{\partial}{\partial \tilde{x}}.
\end{align*}
Evaluated at our point of interest we obtain $\partial_\theta=\partial_{\tilde y}$ and $\partial_\phi=\sin\theta_0\partial_{\tilde x}$. The second-order derivatives can then be obtained using the first-order derivatives. We can immediately evaluate them at the point to get
\begin{align*}
    \partial_\phi^2|_{(r_0,\theta_0,\phi_0)}&=-\sin^2\theta_0\partial_{\tilde r}-\sin\theta_0\cos\theta_0\partial_{\tilde y}+\sin^2\theta_0\partial^2_{\tilde x},\\
    \partial_\theta\partial_\phi|_{(r_0,\theta_0,\phi_0)}&=\cos\theta_0\partial_{\tilde x}+\sin\theta_0\partial_{\tilde x}\partial_{\tilde y},\\
    \partial_\theta^2|_{(r_0,\theta_0,\phi_0)}&=-\partial_{\tilde r}+\partial^2_{\tilde y}.
\end{align*}
Thus, at $(r_0, \theta_0, \phi_0)$,
\begin{gather*}
    \frac{1}{2}\eth_1(\eth_0\psi) = \frac{1}{2}\sin\theta(\partial_\theta+\frac{i}{\sin\theta}\partial_\phi)(\frac{1}{\sin\theta}(\partial_\theta+\frac{1}{\sin\theta}\partial_\phi))\\
    =\frac{\partial^2 \psi}{\partial \theta^2} - \frac{\cos\theta}{\sin\theta} \frac{\partial \psi}{\partial \theta} + \frac{2 i}{\sin\theta} \frac{\partial^2 \psi}{\partial \theta \partial \phi} - \frac{1}{\sin^2\theta} \frac{\partial^2 \psi}{\partial \phi^2} - 2 i \frac{\cos\theta}{\sin^2\theta} \frac{\partial \psi}{\partial \phi}\\
    =\frac{1}{2}(\partial_{\tilde y}^2 - \partial_{\tilde x}^2 + 2i\partial_{\tilde x}\partial_{\tilde y})\psi
    =\gamma_1 + i\gamma_2 = \gamma.
\end{gather*}

\section{Numerical derivative}
\label{sec:derivatives}
The derivatives with respect to cosmological parameters were taken with a central difference formula, i.e.
\begin{equation*}
    f'(x) = \frac{f(x+h) - f(x-h)}{2h} + \mathcal O (h^2).
\end{equation*}
Each change of the cosmological parameters requires a recalculation of the entire cosmology, making it computationally expensive. For the approximation to be accurate, a balance needs to be found between numerical errors for small $h$ and a larger $\mathcal O(h^2)$ error for larger $h$. The $h$ values chosen for each parameter are shown in table \ref{tab:cosmo-params-diff} and are similar to the values used in \cite{Namikawa_2016}\footnote{As confirmed during a conversation with the author.}.

\begin{table}[h!]
    \centering
    \begin{tabular}{|l|c|c|}
    \hline
    Parameter & Fiducial value & Finite difference ($h$) \\
    \hline
    $H$            & $67.4$               & fiducial\,$\times0.1$   \\
    $\Omega_b h^2$ & $0.0223$             & fiducial\,$\times0.1$   \\
    $\Omega_c h^2$ & $0.119$              & fiducial\,$\times0.005$ \\
    $n_s$          & $0.965$              & fiducial\,$\times0.005$ \\
    $A_s$          & $2.13\times10^{-9}$  & fiducial\,$\times0.1$   \\
    $\tau$         & $0.063$              & fiducial\,$\times 0.1$   \\
    $m_\nu$        & $0.06$               & fiducial\,$\times0.1$   \\
    $w_0$          & $-1$                 & $0.03$                  \\
    \hline
    \end{tabular}
    \caption{Fiducial cosmological parameters and their finite-difference steps}
    \label{tab:cosmo-params-diff}
\end{table}

To test the accuracy, we varied $h$ by $\pm 5\%$ and $\pm 10\%$ and compared the relative change in the derivative of the equilateral lensing bispectra and the lensing power spectra. As long as numerical noise doesn't dominate, it can be shown that the relative error in our approximation is approximately 5 times the relative difference that taking $h \rightarrow h(1 \pm 0.1)$ leads to. We thus conclude that, based on the tests conducted, the derivatives are almost always computed with up to one percent error. The exception is the derivative with respect to neutrino masses, which introduces a larger error of around 10 percent.

% \begin{figure}[t]
%     \centering
%     \includegraphics[height=\textheight]{../code/plots/derivative_accuracy.png}
%     \caption{Central difference approximation of derivative of equilateral CMB weak lensing bispectrum (left) and power spectrum (right) for differences as specified in table \ref{tab:cosmo-params-diff} as well as for differences $5\%$ and $10\%$ larger and smaller.}
%     \label{fig:deraccuracy}
% \end{figure}
\pagebreak
\section{$\Lambda$CDM constraints}
\label{sec:lambdacdmconstraints}

We do not expect to use weak lensing surveys to competitively constrain the main $\Lambda$CDM parameters. Despite this, we include these constraints for completeness in figures \ref{fig:paramconstraintsallcmb}, \ref{fig:paramconstraintsallcmbcmbprior}, \ref{fig:paramconstraintsallgal}, \ref{fig:paramconstraintsallgalcmbprior}, and in table \ref{tab:paramconstraintsall}. All survey parameters are the same as in section \ref{sec:results}.
\begin{table}[h!]
    \centering
    \tiny
\begin{tabular}{|l|r|rrr|rrr|rrr|}
    \hline
    \multicolumn{11}{|c|}{weak priors} \\
    \hline
    && \multicolumn{3}{c|}{CMB lensing} & \multicolumn{3}{c|}{Gal. lensing} & \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
    \hline
    Par & prior & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_2}$ & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
    \hline
$H_0$              & 17    & 17    & 16    & 6.5   & 13    & 4.2   & 1.8   & 1.6   & 1.5   & 1.0   \\
$10^3 \Omega_b h^2$ & 0.50  & 0.50  & 0.50  & 0.50  & 0.50  & 0.48  & 0.48  & 0.50  & 0.48  & 0.46  \\
$\Omega_c h^2$     & 0.29  & 0.0070& 0.012 & 0.0063& 0.0095& 0.0049& 0.0041& 0.0050& 0.0044& 0.0033\\
$n_s$              & 0.020 & 0.019 & 0.020 & 0.018 & 0.020 & 0.017 & 0.011 & 0.015 & 0.016 & 0.010 \\
$\tau$             & 0.063 & 0.063 & 0.063 & 0.063 & 0.063 & 0.063 & 0.063 & 0.063 & 0.063 & 0.063 \\
$10^9 A_s$         & 1.0   & 0.36  & 0.91  & 0.30  & 0.53  & 0.10  & 0.10  & 0.081 & 0.098 & 0.063 \\
    \hline
    \multicolumn{11}{|c|}{CMB $T + E$ priors} \\
    \hline
    && \multicolumn{3}{c|}{CMB lensing} & \multicolumn{3}{c|}{Gal. lensing} & \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
    \hline
Par & prior & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_2}$ & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
\hline
$H_0$              & 1.2   & 1.1   & 1.2   & 1.1   & 1.1   & 0.90  & 0.66  & 0.84  & 0.78  & 0.43  \\
$10^3 \Omega_b h^2$ & 0.057 & 0.053 & 0.057 & 0.053 & 0.056 & 0.053 & 0.052 & 0.051 & 0.052 & 0.050 \\
$\Omega_c h^2$     & 0.00083& 0.00076& 0.00081& 0.00076& 0.00064& 0.00068& 0.00053& 0.00048& 0.00061& 0.00043 \\
$n_s$              & 0.0025& 0.0025& 0.0025& 0.0025& 0.0022& 0.0024& 0.0022& 0.0020& 0.0022& 0.0018 \\
$\tau$             & 0.013 & 0.011 & 0.012 & 0.011 & 0.011 & 0.010 & 0.0093& 0.0091& 0.0098& 0.0082 \\
$10^9 A_s$         & 0.055 & 0.044 & 0.049 & 0.043 & 0.045 & 0.041 & 0.037 & 0.036 & 0.039 & 0.032 \\
\hline
\end{tabular}
        \caption{Same as table \ref{tab:paramconstraintstight}, except for $\Lambda$CDM parameters.}
        \label{tab:paramconstraintsall}
\end{table}

\begin{figure}[h!]
    \centering
    \includegraphics[width=\textwidth]{paper/figures/param_constraints_lcdm_cmb_weak_prior.pdf}
    \caption{Same as figure \ref{fig:paramconstraintstightcmb}, except for the standard $\Lambda$CDM parameters.}
    \label{fig:paramconstraintsallcmb}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=\textwidth]{paper/figures/param_constraints_lcdm_cmb_cmb_prior.pdf}
    \caption{Same as figure \ref{fig:paramconstraintstightcmbcmbprior}, except for the standard $\Lambda$CDM parameters.}
    \label{fig:paramconstraintsallcmbcmbprior}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=\textwidth]{paper/figures/param_constraints_lcdm_gal_weak_prior.pdf}
    \caption{Same as figure \ref{fig:paramconstraintstightgal}, except for the standard $\Lambda$CDM parameters.}
    \label{fig:paramconstraintsallgal}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=\textwidth]{paper/figures/param_constraints_lcdm_gal_cmb_prior.pdf}
    \caption{Same as figure \ref{fig:paramconstraintstightgalcmbprior}, except for the standard $\Lambda$CDM parameters.}
    \label{fig:paramconstraintsallgalcmbprior}
\end{figure}

% NOT SUPPOSED TO STAY IN THE PAPER, JUST FOR REFERENCE
% \section{Papers compared to}
% Takada Jain 2005:
% \begin{itemize}
%     \item $ 50 < l < 3000$
%     \item powerspectrum + bispectrum
%     \item galaxy lensing
%     \item $\epsilon = 0.4$, $n_g = 100$, $f_{sky}=0.1$
%     \item derivative step: 5 percent.
%     \item priors:
%     \begin{gather*}
%         \sigma(\ln \Omega_B h^2) = 0.01\\
%         \sigma(n) = 0.008\\
%         \sigma(H_0) = 13
%         \sigma(m_\nu) = \infty
%     \end{gather*}
%     \item constraints obtained by paper:
%     \begin{table}[h!]
%         \centering
%         \begin{tabular}{lccc}
%           \hline
%           Quantity                & PS   & Bisp & PS+Bisp \\
%           \hline
%           $\sigma(\Omega_{\rm de})$ & 0.14 & 0.13 & 0.023   \\
%           $\sigma(\sigma_8)$        & 0.12 & 0.12 & 0.023   \\
%           $\sigma(w_0)$             & 4.0  & 3.6  & 1.1     \\
%           $\sigma(w_a)$             & 20   & 15   & 5.1     \\
%           \hline
%         \end{tabular}
%       \end{table}
%       \item constraints we got:
%       \begin{table}[h!]
%         \centering
%         \begin{tabular}{lccc}
%           \hline
%           Quantity                & PS   & Bisp & PS+Bisp \\
%           \hline
%           $\sigma(\sigma_8)$        & 0.09 & 0.09 & 0.05   \\
%           $\sigma(w_0)$             & 0.37  & 0.37  & 0.21     \\
%           \hline
%         \end{tabular}
%       \end{table}
% \end{itemize}

% Planck VII:
% \begin{itemize}
%     \item $ 8 < l < 400$
%     \item powerspectrum only
%     \item cmb lensing
%     \item noise: 5 beamwidth, $\Delta_T = 30$, $\Delta_P = 52$.
%     \item derivative step: ??
%     \item priors:
%     \begin{gather*}
%         \sigma(\Omega_b h^2) = 0.0005\\
%         \sigma(\Omega_c h^2) = [0.01, 0.99]\\
%         \sigma(n_s) = 0.02\\
%         \sigma(10^10A_s) = [1.91, 3.91]\\
%         \sigma(H_0) = [0.4, 1.0]\\
%         \sigma(m_\nu) = \infty\\
%         \sigma(w_0) = \infty
%     \end{gather*}
%     \item constraints obtained by paper:
%     \begin{table}[h!]
%         \centering
%         \begin{tabular}{lc}
%           \hline
%           Quantity & PS  \\
%           \hline
%           $\sigma(\sigma_8\Omega_m^{0.25})$ & 0.020 \\
%           \hline
%         \end{tabular}
%       \end{table}
%       \item constrain that we obtained: 0.0275
%       \item notes: actually different beamwidth for polarization and temperature
% \end{itemize}

\end{document}